---
title: "Figure 1"
author: "M. Pohly"
date: "27/04/2023"
output: html_document
---
Last updated on `r Sys.Date()`
This figure will provide an overview over the data sets, patients and compounds. 

#### Load libraries
```{r warning=FALSE, message=FALSE}
library(readxl)
library(tidyverse)
library(DrugScreenExplorer)
library(RColorBrewer)
library(ComplexHeatmap)
library(ggvenn)
library(ggtext)
library(parallel)
library(ggbeeswarm)
library(gridExtra)
library(circlize)
library(ggrepel)
```

#### Define variables
```{r}
opt <- list()
opt$drugscreen <- "data/submission/drugScreens_pseudo.RDS"
opt$druganno <- "misc/drugList_suppl.xlsx"
opt$plot <- "plots/Fig1/"

## Set theme
lgd <-  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5), 
        text = element_text(size = 15),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA), 
        legend.background = element_rect(fill='transparent',colour = NA), 
        strip.background = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
```

#### Load drug screen data and annotations
```{r}
drugList <- readRDS(opt$drugscreen)
drugAnno <- read_excel(opt$druganno) %>%
  dplyr::select(-Supplier) %>% unique()
```

#### Fig 1B
#### Bar plots of drug composition
```{r}
drugAnno.red <- drugAnno %>%
  mutate(Pathway = ifelse(Pathway %in% c("HDAC", "HDM", "BET", "HMT", "HAT", "HMT, MYC", "NFKB, HDAC, DNA Damage"), "Epigenome", Pathway)) %>%
  mutate(Pathway = ifelse(Pathway %in% c("Apoptosis", "Necroptosis", "Autophagy"), "Prg. Cell Death", Pathway)) %>%
  mutate(Pathway = ifelse(Pathway %in% c("DNA Damage", "DNA Damage, NFKB", "Chemo"), "DNA Damage", Pathway)) %>%
  mutate(Pathway = ifelse(Pathway %in% c("EGFR", "Prg. Cell Death", "BRAF", "Chemo", "Epigenome", "JAK-STAT", "DNA Damage", 
                                         "Cell Cycle", "BCR/TCR", "MEK", "PI3K-AKT-mTOR", "MAPK", "Metabolism", "ROS"), Pathway, "Other")) %>%
  group_by(Screen) %>% mutate(TotalDrugs = length(Drug)) %>% ungroup()


## ScreenA
pathOrder <- c("PI3K-AKT-mTOR", "Epigenome", "Cell Cycle", "Metabolism", "DNA Damage",
               "BCR/TCR", "MAPK", "ROS", "JAK-STAT", "Prg. Cell Death", "BRAF", 
               "EGFR", "MEK", "Other")

colTab <- c("#1E88E5", "#009688", "#FF8F00", "#E53935", "#9575CD", "#4FC3F7", #"#4FC3F7",
                     "#81C784", "#F06292", "#EF9A9A", "#C5E1A5", "#B3E5FC", "#FDD835" , "#FFF9C4", "#1A237E")

geomText <- drugAnno.red %>% 
  filter(Screen != "Screen E") %>% 
  select(Screen, TotalDrugs) %>% unique()

p <- drugAnno.red %>% filter(Screen != "Screen E") %>%
  group_by(Screen) %>%
  dplyr::count(Pathway, TotalDrugs) %>%
  mutate(percentage = n / TotalDrugs) %>%
#  arrange(desc(n)) %>%
  mutate(Pathway = factor(Pathway, levels = pathOrder), 
         Screen = factor(Screen, levels = c("Screen A", "Screen B", "Screen C", "Screen D"))) %>%
  #mutate(PathwayPer = paste(Pathway, " (", n, ")", sep = "")) %>% 
  arrange(Pathway) %>%
  #mutate(Pathway = factor(PathwayPer, levels = PathwayPer)) %>%
  ggplot(aes(x = Screen, y = percentage)) +
  geom_bar(stat = "identity", colour = "black", aes(fill = Pathway)) +
  geom_text(aes(label = TotalDrugs, x = Screen), y = 1.025, data = geomText, size = 5) +
  xlab("") + ylab("Fraction of the Drug Library") + ggtitle("Drug Libraries") +
  #coord_polar("y", start=0, direction = -1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5), 
        text = element_text(size = 17.5),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA), 
        legend.background = element_rect(fill='transparent',colour = NA), 
        strip.background = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(linewidth = 1)) +
  scale_fill_manual(values = colTab)
p

ggsave(plot = p, filename = paste0(opt$plot, "druglibrary_bar.png"), height = 15, 
       width = 20, units = "cm")

```

#### Bar chart of screen composition
```{r}
plotTab <- lapply(c("ScreenA", "ScreenB", "ScreenC", "ScreenD"), function(x) {
  screenData <- drugList[[x]]
  
  screenData %>% filter(diagnosis %in% c("CLL", "MCL", "T-PLL")) %>%
    select(patientID, diagnosis) %>%
    unique() %>%
    dplyr::count(diagnosis) %>%
    mutate(totalPatients = sum(n), 
           rel = n / totalPatients,
           Screen = x)
}) %>% bind_rows() %>%
  mutate(Screen = str_replace(string = Screen, pattern = "Screen", replacement = "Screen "))


p <- plotTab %>%
  mutate(Screen = factor(Screen, levels = c("Screen A", "Screen B", "Screen C", "Screen D")), 
         Diagnosis = diagnosis) %>%
  ggplot(aes(x = Screen, y = rel)) +
  geom_bar(stat = "identity", colour = "black", aes(fill = Diagnosis)) +
  geom_text(aes(label = totalPatients, x = Screen), y = 1.025, size = 5, data = unique(select(plotTab, Screen, totalPatients))) +
  xlab("") + ylab("Fraction of the Total Patients") + ggtitle("Patients per Screen") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5), 
        text = element_text(size = 17.5),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA), 
        legend.background = element_rect(fill='transparent',colour = NA), 
        strip.background = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(linewidth = 1)) +
  scale_fill_manual(values = c("MCL" = "#F44336", "T-PLL" = "#64B5F6", "CLL" = "#FFD45F"))
p

ggsave(plot = p, filename = paste0(opt$plot, "patients_bar.png"), height = 15, 
       width = 19, units = "cm")

```

##### Piechart - Screen A
```{r fig.height= 5, fig.width=5}
## Make annotation more uniform
drugAnno.red <- drugAnno %>%
  mutate(Pathway = ifelse(Pathway %in% c("HDAC", "HDM", "BET", "HMT", "HAT", "HMT, MYC", "NFKB, HDAC, DNA Damage"), "Epigenome", Pathway)) %>%
  mutate(Pathway = ifelse(Pathway %in% c("Apoptosis", "Necroptosis", "Autophagy"), "Prg. Cell Death", Pathway)) %>%
  mutate(Pathway = ifelse(Pathway %in% c("DNA Damage", "DNA Damage, NFKB", "Chemo"), "DNA Damage", Pathway)) %>%
  mutate(Pathway = ifelse(Pathway %in% c("EGFR", "Prg. Cell Death", "BRAF", "Chemo", "Epigenome", "JAK-STAT", "DNA Damage",
                                         "Cell Cycle", "BCR/TCR", "MEK", "PI3K-AKT-mTOR", "MAPK", "Metabolism", "ROS"), Pathway, "Other")) %>%
  group_by(Screen) %>% mutate(TotalDrugs = length(Drug)) %>% ungroup()

## ScreenA
pathOrder <- c("PI3K-AKT-mTOR", "Epigenome", "Cell Cycle", "Metabolism", "DNA Damage",
               "BCR/TCR", "MAPK", "ROS", "JAK-STAT", "Prg. Cell Death", "BRAF",
               "EGFR", "MEK", "Other")

colTab <- c("#1E88E5", "#009688", "#FF8F00", "#E53935", "#9575CD", "#4FC3F7", #"#4FC3F7",
                     "#81C784", "#F06292", "#EF9A9A", "#C5E1A5", "#B3E5FC", "#FDD835" , "#FFF9C4", "#1A237E")

p <- drugAnno.red %>% filter(Screen == "Screen A") %>%
  dplyr::count(Pathway, TotalDrugs) %>%
  mutate(percentage = n / TotalDrugs) %>%
#  arrange(desc(n)) %>%
  mutate(Pathway = factor(Pathway, levels = pathOrder)) %>%
  mutate(PathwayPer = paste(Pathway, " (", n, ")", sep = "")) %>%
  arrange(Pathway) %>%
  mutate(Pathway = factor(PathwayPer, levels = PathwayPer)) %>%
  ggplot(aes(x = "", y = percentage, fill = Pathway)) +
  geom_bar(stat = "identity", colour = "black") +
  #ggtitle("Drug Library") +
  ggtitle("") +
  coord_polar("y", start=0, direction = -1) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size = 12.5, face = "bold"),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        legend.background = element_rect(fill='transparent',colour = NA),
        strip.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  scale_fill_manual(values = colTab)
p

ggsave(plot = p, filename = paste0(opt$plot, "ScreenA_piechar.png"), height = 5,
       width = 5)
```

##### Piechart - ScreenB
```{r fig.height= 5, fig.width=5}
## ScreenB
p <- drugAnno.red %>% filter(Screen == "Screen B") %>%
  dplyr::count(Pathway, TotalDrugs) %>%
  mutate(percentage = n / TotalDrugs) %>%
  #  arrange(desc(n)) %>%
  mutate(Pathway = factor(Pathway, levels = pathOrder)) %>%
  mutate(PathwayPer = paste(Pathway, " (", n, ")", sep = "")) %>%
  arrange(Pathway) %>%
  mutate(Pathway = factor(PathwayPer, levels = PathwayPer)) %>%
  ggplot(aes(x = "", y = percentage, fill = Pathway)) +
  geom_bar(stat = "identity", colour = "black") +
  #ggtitle("Drug Library") +
  ggtitle("") +
  coord_polar("y", start=0, direction = -1) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size = 12.5, face = "bold"),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        legend.background = element_rect(fill='transparent',colour = NA),
        strip.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  scale_fill_manual(values = colTab)
p

ggsave(plot = p, filename = paste0(opt$plot, "ScreenB_piechar.png"), height = 5,
       width = 5)
```

##### Piechart - ScreenC
```{r fig.height= 5, fig.width=5}
## ScreenC
p <- drugAnno.red %>% filter(Screen == "Screen C") %>%
  dplyr::count(Pathway, TotalDrugs) %>%
  mutate(percentage = n / TotalDrugs) %>%
  #  arrange(desc(n)) %>%
  mutate(Pathway = factor(Pathway, levels = pathOrder)) %>%
  mutate(PathwayPer = paste(Pathway, " (", n, ")", sep = "")) %>%
  arrange(Pathway) %>%
  mutate(Pathway = factor(PathwayPer, levels = PathwayPer)) %>%
  ggplot(aes(x = "", y = percentage, fill = Pathway)) +
  geom_bar(stat = "identity", colour = "black") +
  #ggtitle("Drug Library") +
  ggtitle("") +
  coord_polar("y", start=0, direction = -1) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size = 12.5, face = "bold"),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        legend.background = element_rect(fill='transparent',colour = NA),
        strip.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  scale_fill_manual(values = colTab)
p

ggsave(plot = p, filename = paste0(opt$plot, "ScreenC_piechar.png"), height = 5,
         width = 5)
```

##### Piechart - ScreenD
```{r fig.height= 5, fig.width=5}
## ScreenD
colTab <- c("#1E88E5", "#009688", "#FF8F00", "#9575CD", "#4FC3F7", #"#4FC3F7",
                     "#81C784", "#F06292", "#EF9A9A", "#C5E1A5", "#B3E5FC", "#FDD835" , "#FFF9C4", "#1A237E")

p <- drugAnno.red %>% filter(Screen == "Screen D") %>%
  dplyr::count(Pathway, TotalDrugs) %>%
  mutate(percentage = n / TotalDrugs) %>%
  #  arrange(desc(n)) %>%
  mutate(Pathway = factor(Pathway, levels = pathOrder)) %>%
  mutate(PathwayPer = paste(Pathway, " (", n, ")", sep = "")) %>%
  arrange(Pathway) %>%
  mutate(Pathway = factor(PathwayPer, levels = PathwayPer)) %>%
  ggplot(aes(x = "", y = percentage, fill = Pathway)) +
  geom_bar(stat = "identity", colour = "black") +
  #ggtitle("Drug Library") +
  ggtitle("") +
  coord_polar("y", start=0, direction = -1) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size = 12.5, face = "bold"),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        legend.background = element_rect(fill='transparent',colour = NA),
        strip.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  scale_fill_manual(values = colTab)
p

ggsave(plot = p, filename = paste0(opt$plot, "ScreenD_piechar.png"), height = 5,
       width = 5)

```

##### Vendiagram - Patients
```{r fig.height= 8, fig.width=8}
## Get screen data
ScreenA <- drugList[["ScreenA"]]
ScreenB <- drugList[["ScreenB"]]
ScreenC <- drugList[["ScreenC"]]
ScreenD <- drugList[["ScreenD"]]

## Adjust info and get into a single data.frame
ScreenA.df <- ScreenA %>% filter(diagnosis == "T-PLL") %>%
  select(patientID) %>% unique() %>%
  mutate(screen = "Screen A")

ScreenB.df <- ScreenB %>% filter(diagnosis == "T-PLL") %>%
  select(patientID) %>% unique() %>%
  mutate(screen = "Screen B")

ScreenC.df <- ScreenC %>% filter(diagnosis == "T-PLL") %>%
  select(patientID) %>% unique() %>%
  mutate(screen = "Screen C")

ScreenD.df <- ScreenD %>% filter(diagnosis == "T-PLL") %>%
  select(patientID) %>% unique() %>%
  mutate(screen = "Screen D")

patDat <- rbind(ScreenA.df, ScreenB.df, ScreenC.df, ScreenD.df) %>%
  pivot_wider(names_from = "screen", values_from = "screen")

## Prepare list for venndiaggram
x <- list(
  `Screen B` = patDat %>% filter(`Screen B` == "Screen B") %>% select(patientID) %>%
    unlist() %>% as.vector(), 
  `Screen A` = patDat %>% filter(`Screen A` == "Screen A") %>% select(patientID) %>%
    unlist() %>% as.vector(), 
  `Screen C` = patDat %>% filter(`Screen C` == "Screen C") %>% select(patientID) %>%
    unlist() %>% as.vector(),
  `Screen D` = patDat %>% filter(`Screen D` == "Screen D") %>% select(patientID) %>%
    unlist() %>% as.vector())

## Make venndiagram with labels
p <- ggvenn(x, fill_color = c( "#EF9A9A", "#D1C4E9", "#B3E5FC", "#A5D6A7"),
            stroke_size = 0.75, set_name_size = 11, text_size = 11, show_percentage = FALSE) + 
  theme(plot.title = element_text(hjust=0.5), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA))
p


## Make venndiagram with labels
p <- ggvenn(x, fill_color = c( "#EF9A9A", "#D1C4E9", "#B3E5FC", "#A5D6A7"),
            stroke_size = 0.75, set_name_size = 0, text_size = 11, show_percentage = FALSE) + 
theme(plot.title = element_text(hjust=0.5), 
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "transparent",colour = NA),
      plot.background = element_rect(fill = "transparent",colour = NA))
  
ggsave(plot = p, filename = paste0(opt$plot, "patientoverlap.png"), 
       height = 8, width = 8)
```

##### Vendiagram - Compounds
```{r fig.height= 8, fig.width=8}
## Get into the correct format
drugAnno.venn <- drugAnno %>% select(Drug, Screen) %>%
  pivot_wider(names_from = "Screen", values_from = "Screen") %>%
  unique()

# Drug overlap
x <- list(
  `Screen B` = drugAnno.venn %>% filter(`Screen B` == "Screen B") %>% select(Drug) %>%
    unlist() %>% as.vector(), 
  `Screen A` = drugAnno.venn %>% filter(`Screen A` == "Screen A") %>% select(Drug) %>%
    unlist() %>% as.vector(), 
  `Screen C` = drugAnno.venn %>% filter(`Screen C` == "Screen C") %>% select(Drug) %>%
    unlist() %>% as.vector(),
  `Screen D` = drugAnno.venn %>% filter(`Screen D` == "Screen D") %>% select(Drug) %>%
    unlist() %>% as.vector()
)

## Version with text:
p <- ggvenn(x, fill_color = c( "#EF9A9A", "#D1C4E9", "#B3E5FC", "#A5D6A7"),
            stroke_size = 0.75, set_name_size = 11, text_size = 11, show_percentage = FALSE) + 
  theme(plot.title = element_text(hjust=0.5), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA))
p


## Version without text
p <- ggvenn(x, fill_color = c( "#EF9A9A", "#D1C4E9", "#B3E5FC", "#A5D6A7"),
            stroke_size = 0.75, set_name_size = 0, text_size = 11, show_percentage = FALSE) + 
theme(plot.title = element_text(hjust=0.5), 
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      panel.background = element_rect(fill = "transparent",colour = NA),
      plot.background = element_rect(fill = "transparent",colour = NA))
  
ggsave(plot = p, filename = paste0(opt$plot, "drugoverlap.png"), 
       height = 8, width = 8)

```

#### Fig 1C
##### Ranked toxicity - ScreenA
```{r message=FALSE, fig.height= 6, fig.width=15}
ScreenA <- drugList[["ScreenA"]] # get ScreenA data

## index concentrations
ScreenA.ind <- mclapply(unique(ScreenA$name), mc.cores = 4, function(x) {
  message("Running ", x)
  ScreenA.x <- dplyr::filter(ScreenA, name == x) %>% 
    mutate(conc = as.numeric(conc))
  subTab <- ScreenA.x %>%
    dplyr::select(name, conc, concIndex) %>% arrange(desc(conc)) %>% unique()
 subTab$Index <- seq(1:nrow(subTab))
  message("Adding ", nrow(subTab), " indices")
  subTab %>% left_join(ScreenA.x, by = c("name", "conc", "concIndex"))
}) %>% bind_rows()

## compute the mean druf effect per patient x drug
viabTab <- ScreenA.ind %>% dplyr::filter(!is.na(name), !is.na(diagnosis)) %>%
  dplyr::filter(!name %in% c("PBS", "DMSO")) %>%
  group_by(patientID, name, diagnosis) %>%
  mutate(value = mean(viab.auc, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(dplyr::select(drugAnno, Drug, Target, Pathway), 
            by = c("name" = "Drug"))

topN <- viabTab$name %>% unique() %>% length() # define the number of drugs to plot
colorCode <- rev(brewer.pal(6,"Blues"))[1:6] # define colour scheme

## Subset to T-PLL patients
viabTab <- dplyr::filter(viabTab, diagnosis == "T-PLL") %>%
  mutate(Index = factor(Index)) %>%
  dplyr::select(patientID, diagnosis, name, Target, Pathway, viab.auc, value, viab, concIndex, Index) %>%
  unique()

## reshuffle before plotting
set.seed(100)
viabTab <- viabTab[sample(1:nrow(viabTab),nrow(viabTab)), ]

## Make ranked toxicity plot
p <- lapply(unique(viabTab$diagnosis), function(n) {
  eachTab <- dplyr::filter(viabTab, diagnosis == n)
  drugOrder <- group_by(eachTab, name) %>%
    summarise(medVal = median(value, na.rm = TRUE)) %>%
    arrange(medVal)
  drugOrder <- drugOrder[c(seq(1,topN)), ]
  eachTab <- dplyr::filter(eachTab, name %in% drugOrder$name) %>%
    mutate(name = factor(name, levels = drugOrder$name), 
           concIndex = as.factor(concIndex)) %>%
    group_by(name) %>% mutate(medVal = median(value, na.rm = TRUE)) %>%
    ungroup()
  drugOrder$rank <- seq(1:nrow(drugOrder))
  eachTab <- left_join(eachTab, dplyr::select(drugOrder, name, rank), by = "name")
  eachTab <- eachTab %>%
    group_by(name#, concIndex
             ) %>%
    mutate(viab.auc = mean(viab, na.rm = TRUE))
  p <- ggplot(eachTab, aes(x=rank,y = viab.auc#, fill = Index
                           )) +
    geom_jitter(alpha=0.8, shape = 21, size = 3, height = 0, fill = "#64B5F6") +
    #scale_color_manual(values = colorCode) + 
    scale_x_continuous(expand = c(0.005, 0), labels = c(100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200, 1300, 1400, 1500, 1600, 1700, 1800, 1900, 2000, 2100, 2200, 2300, 2400, 2500), 
                       breaks = c(100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200, 1300, 1400, 1500, 1600, 1700, 1800, 1900, 2000, 2100, 2200, 2300, 2400, 2500)) +
    ylab("Viability") + ylim(c(0,1.2)) + 
    xlab("Rank") + #ggtitle("ScreenA - Ranked Efficacy") +
    theme_classic() +
    geom_hline(yintercept = 1, linetype = "dashed") +
    theme(text = element_text(size = 17.5), #axis.text.x=element_blank(),
      #axis.ticks.x=element_blank(), 
      #axis.ticks = element_blank(), 
      legend.key = element_blank(),
      axis.line = element_line(linewidth = 0.75),
      #legend.position = "none",
      plot.title = element_text(hjust=0.5), 
      # plot.margin = unit(c(0.5, 0.5, 0.0, 0), "cm"), 
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      #legend.direction="horizontal",
      #legend.position = c(0.2, 0.25),
      panel.background = element_rect(fill = "transparent",colour = NA),
      plot.background = element_rect(fill = "transparent",colour = NA), 
      legend.background = element_rect(fill='transparent')) +
    guides(colour = guide_legend(override.aes = list(size=3.5), title="Concentration\nIndex"))
}) 
p

ggsave(plot = p[[1]], filename = paste0(opt$plot, "/ScreenA_overview_low.png"),
       height = 6.5, width = 20)
```

##### Drug-Drug correlation - ScreenA
```{r fig.height= 12.5, fig.width=12.5}

# viabMat <- filter(ScreenA, diagnosis == "T-PLL") %>%
#   group_by(patientID, name) %>% summarise(viab = mean(viab, na.rm = TRUE)) %>%
#   spread(key = patientID, value = "viab") %>% data.frame() %>%
#   column_to_rownames("name") %>% as.matrix()
# 
# viabMat <- viabMat[complete.cases(viabMat),]
# 
# ## Define colour-scheme
# colorList <- c(colorRampPalette(c("lightseagreen", "white"))(20),
#                colorRampPalette(c("white"))(10),
#                colorRampPalette(c("white","deeppink"))(20))
# 
# # corMat <- cor(t(viabMat), method = "spearman")
# corMat <- cor(t(viabMat), method = "pearson")
# 
# p4 <- pheatmap(corMat, color = colorList, 
#                show_rownames = FALSE, show_colnames = FALSE, legend = FALSE,
#                breaks = seq(-1,1,length.out = 50), clustering_method = "ward.D2", name = "R^2", 
#                )
# p4

# if(opt$saveplots == TRUE) {
#   png(file="plots/fig1/drug-drug-matrix.png", 
#     height = 25, width = 25, res = 600, units = "cm", bg = "transparent")
#   draw(p4, background = "transparent")
#   dev.off()
# }

```

##### TPLL drug response - ScreenA
```{r fig.height= 5.5, fig.width=5.5}

# tTest <- function(value, type) {
#   res <- t.test(value ~ factor(type), equal.var = TRUE, na.action = na.exclude)
#   tibble(p = res$p.value,
#          diff = res$estimate[[2]] - res$estimate[[1]])
# }
# 
# compDiag <- "CLL"
# diagSelected <- c(compDiag, "T-PLL")
# 
# screenDat <- ScreenA
# 
# ## Compute average toxicity and perform pairwise t-test
# pTab <- lapply(diagSelected[diagSelected != c(compDiag)], function(diag) {
#   print(diag)
#   testTab <- filter(screenDat, diagnosis %in% c(compDiag, diag),
#                     !name %in% c("DMSO", "empty")) %>%
#     #filter(name %in% c("Venetoclax")) %>%
#     mutate(diagnosis = factor(diagnosis, levels = c(compDiag, diag)), 
#            viab = viab.auc) %>%
#     select(patientID, diagnosis, name, viab) %>% filter(!is.na(viab)) %>% unique()
#   res <- group_by(testTab, name) %>% do(tTest(.$viab, .$diagnosis))
#   res$diagnosis <- diag
#   res
# }) %>% bind_rows() %>% ungroup() %>% mutate(p.adj = p.adjust(p, method = "BH"))
# 
# plotTab <- pTab %>% mutate(p = ifelse(p < 1e-12, 1e-12, p)) %>%
#   mutate(pSign = -log10(p)*sign(diff)) %>%
#   ungroup()
# 
# plotTab$diffCol <- "ns"
# plotTab[plotTab[, "p"] < 0.05 & plotTab[, "diff"] < -0.05, "diffCol"] <- "higher_sen"
# plotTab[plotTab[, "p"] < 0.05 & plotTab[, "diff"] > 0.05, "diffCol"] <- "lower_sen"
# 
# p1 <- plotTab %>%
#   ggplot(aes(x = diff, y = -log10(p))) +
#   geom_point(shape = 21, size = 2.5, aes(fill = diffCol)) +
#   geom_hline(yintercept = 1.3, linetype = "dashed") +
#   geom_vline(xintercept = 0.05, linetype = "dashed") +
#   geom_vline(xintercept = -0.05, linetype = "dashed") +
#   scale_x_reverse(limits = c(0.5, -0.5)) +
#   xlab("Viability Difference") + ylab("-log10(p.value)") +
#   ggtitle("") +
#   theme_classic() + 
#   theme(plot.title = element_text(hjust = 0.5, size = 20), 
#         axis.title = element_text(size = 17.5),
#         axis.text = element_text(size = 15),
#         axis.line = element_line(linewidth = 0.75),
#         legend.position = "none",
#         legend.key = element_blank(),
#         panel.grid.major = element_blank(), 
#         panel.grid.minor = element_blank(),
#         panel.background = element_rect(fill = "transparent",colour = NA),
#         plot.background = element_rect(fill = "transparent",colour = NA),
#         legend.background = element_rect(fill='transparent')) +
#   scale_fill_manual(values = c("higher_sen" = "#1976D2", "lower_sen" = "#F44336", "ns" = "grey"))
# p1
# 
# if(opt$saveplots == TRUE) {
#   ggsave(plot = p1, file = "plots/fig1/vol_ScreenA.png", height = 5.5, width = 5.5)
# }


```

##### Piechart - Validation screen
```{r fig.height= 5.5, fig.width=5.5}
## Validation
colTab <- c("#1E88E5", "#009688", "#FF8F00", "#9575CD", "#4FC3F7", #"#4FC3F7",
                     "#81C784", "#EF9A9A", "#C5E1A5", "#FDD835" , "#FFF9C4", "#1A237E")
                     
p <- drugAnno.red %>% filter(Screen == "Screen E") %>%
  dplyr::count(Pathway, TotalDrugs) %>%
  mutate(percentage = n / TotalDrugs) %>%
  #  arrange(desc(n)) %>%
  mutate(Pathway = factor(Pathway, levels = pathOrder)) %>%
  mutate(PathwayPer = paste(Pathway, " (", n, ")", sep = "")) %>% 
  arrange(Pathway) %>%
  mutate(Pathway = factor(PathwayPer, levels = PathwayPer)) %>%
  ggplot(aes(x = "", y = percentage, fill = Pathway)) +
  geom_bar(stat = "identity", colour = "black") +
  #ggtitle("Drug Library") +
  ggtitle("") +
  coord_polar("y", start=0, direction = -1) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5), 
        text = element_text(size = 12.5, face = "bold"),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA), 
        legend.background = element_rect(fill='transparent',colour = NA), 
        strip.background = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  scale_fill_manual(values = colTab)
p

ggsave(plot = p, filename = paste0(opt$plot, "validation_piechar.png"), height = 5, 
       width = 5)
```


#### BH3 mimetics
```{r fig.height= 6.25, fig.width=7.25}
ScreenB <- drugList[["ScreenB"]] # get ScreenB data

viabTab <- ScreenB %>% dplyr::filter(!name %in% c("PBS", "DMSO")) %>%
  group_by(patientID, name, diagnosis) %>%
  mutate(value = mean(viab.auc, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(dplyr::select(drugAnno, Drug, Target, Pathway), 
            by = c("name" = "Drug")) %>%
  dplyr::select(patientID, sampleID, diagnosis, name, Target, Pathway, concIndex, 
         viab, value, viab.auc) %>% unique()

## Subset to BH3-mimetics in T-PLL 
bh3 <- c("Venetoclax", "Navitoclax", "Obatoclax mesylate", "Sabutoclax", "TW-37")
testTab <- viabTab %>% 
  dplyr::filter(diagnosis == "T-PLL", name %in% bh3) %>%
  dplyr::select(patientID, name, viab.auc) %>% unique() %>%
  mutate(name = factor(name, levels = bh3))

fit <- aov(viab.auc ~ name, testTab)
summary(fit)

## Compute p-value
lapply(bh3, function(x) {
  if(x == "Venetoclax") {
    fit.sc <- multcomp::glht(fit, linfct = multcomp::mcp(name = c(4, -1, -1, -1, -1))) # is venetoclax different from the rest?
    }
  if(x == "Navitoclax") {
    fit.sc <- multcomp::glht(fit, linfct = multcomp::mcp(name = c(4, -4, 0, 0, 0)))
    }
  if(x == "Obatoclax mesylate") {
    fit.sc <- multcomp::glht(fit, linfct = multcomp::mcp(name = c(4, 0, -4, 0, 0)))
    }
  if(x == "Sabutoclax") {
    fit.sc <- multcomp::glht(fit, linfct = multcomp::mcp(name = c(4, 0, 0, -4, 0)))
    }
  if(x == "TW-37") {
    fit.sc <- multcomp::glht(fit, linfct = multcomp::mcp(name = c(4, 0, 0, 0, -4)))
    }
  p.adj <- summary(fit.sc, test = multcomp::adjusted("BH"))$test$pvalues
  df <- data.frame(name = x, p.adj = p.adj)
}) %>% bind_rows()

## Get bcl2 affinity from references
bcl2bind <- data.frame(name = c("Venetoclax", "Navitoclax", "Obatoclax mesylate", 
                                "Sabutoclax", "TW-37"), 
                       bcl2aff = c(0.01, 1, 220, 310, 290)) # nM ScreenC
testTab <- testTab %>% left_join(bcl2bind, by = "name")

# fit <- lm(viab.auc ~ bcl2aff, testTab)
# summary(fit)

pVen <- viabTab %>%
  dplyr::filter(name %in% c("Venetoclax", "Navitoclax", "Obatoclax mesylate", 
                     "Sabutoclax", "TW-37"), diagnosis == "T-PLL") %>%
  left_join(bcl2bind, by = "name") %>%
  mutate(name = str_replace(name, "Obatoclax mesylate", "Obatoclax")) %>%
  mutate(name = factor(name, levels = c("Venetoclax", "Navitoclax", "Obatoclax", 
                "Sabutoclax", "TW-37"))) %>%
  dplyr::select(name, patientID, value, bcl2aff) %>% unique() %>%
  ggplot(aes(x = name, y = value, fill = log(bcl2aff))) +
  geom_boxplot(alpha = 0.75) +
  geom_beeswarm(size = 3.5, shape = 21, cex = 3, alpha = 0.75) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  xlab("") + ylab("Mean Viability") +
  ggtitle("Screen B - Diff. Activity of BH3-Mimetics") +
  theme_classic() +
  theme(text = element_text(size = 17.5),
    plot.title = element_text(hjust=0.5), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 20),
    axis.text.y = element_text(size = 20),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
    axis.title = element_text(size = 22.5),
    legend.position = c(0.875, 0.25),
    axis.line = element_line(linewidth = 1),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA), 
        legend.background = element_rect(fill='transparent')) +
  scale_fill_continuous("Log2(Ki Bcl-2)")
pVen

ggsave(plot = pVen, filename = paste0(opt$plot, "diffVen.png"), 
       height = 6.5, width = 7.25)

```

### Dose-response fludarabine
```{r fig.height= 5, fig.width=7}
ScreenC <- drugList[["ScreenC"]] ## get ScreenC data

drugSub <- c("Fludarabine")

concInter <- ScreenC %>%
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  dplyr::select(conc) %>% unique() %>% unlist() %>% as.vector()

p1 <- ScreenC %>% 
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  ggplot(aes(x = conc, y = viab, group = sampleID)) +
  geom_line(size = 0.25) + 
  geom_point(size = 3.75, shape = 21, fill = "#64B5F6") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(limits = c(0, 1.2)) +
  scale_x_log10(labels = concInter, 
                breaks = concInter) +
  xlab("Concentration (µM)") + ylab("Viability") + 
  theme_classic() + ggtitle(paste("Screen C", drugSub, sep = " - ")) +
  theme(text = element_text(size = 17.5), 
        legend.key = element_blank(),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 17.5),
        axis.line = element_line(linewidth = 0.75),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA))
p1

ggsave(plot = p1, filename = paste0(opt$plot, drugSub, ".png"), 
       height = 5, width = 6)
```



#### Dose-response nutlin-3a
```{r fig.height= 5, fig.width=7}
ScreenC <- drugList[["ScreenC"]] ## get ScreenC data

drugSub <- c("Nutlin-3a")

concInter <- ScreenC %>%
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  dplyr::select(conc) %>% unique() %>% unlist() %>% as.vector()

p1 <- ScreenC %>% 
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  ggplot(aes(x = conc, y = viab, group = sampleID)) +
  geom_line(size = 0.25) + 
  geom_point(size = 3.75, shape = 21, fill = "#64B5F6") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(limits = c(0, 1.2)) +
  scale_x_log10(labels = concInter, 
                breaks = concInter) +
  xlab("Concentration (µM)") + ylab("Viability") + 
  theme_classic() + ggtitle(paste("Screen C", drugSub, sep = " - ")) +
  theme(text = element_text(size = 17.5), 
        legend.key = element_blank(),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 17.5),
        axis.line = element_line(linewidth = 0.75),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA))
p1

ggsave(plot = p1, filename = paste0(opt$plot, drugSub, ".png"), 
       height = 5, width = 6)

```

#### Dose-response birinapant
```{r fig.height= 5, fig.width=8}
drugSub <- "Birinapant"

concInter <- ScreenB %>%
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  dplyr::select(conc) %>% unique() %>% unlist() %>% 
  as.vector()
concInter[1:2] <- round(concInter[1:2], 1)
concInter[3:6] <- round(concInter[3:6], 2)
concInter[7:9] <- round(concInter[7:9], 3)

p1 <- ScreenB %>% 
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  ggplot(aes(x = conc, y = viab, group = sampleID)) +
  geom_line(size = 0.25) + 
  geom_point(size = 3.75, shape = 21, fill = "#64B5F6") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(limits = c(0, 1.2)) +
  scale_x_log10(labels = concInter, 
                breaks = concInter) +
  xlab("Concentration (µM)") + ylab("Viability") + 
  theme_classic() + ggtitle(paste("Screen B", drugSub, sep = " - ")) +
  theme(text = element_text(size = 17.5), 
        legend.key = element_blank(),
        plot.title = element_text(hjust=0.5),
        # axis.text.x = element_text(hjust = 1, angle = 45),
        axis.text = element_text(size = 17.5),
        axis.line = element_line(linewidth = 0.75),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA), 
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
p1

ggsave(plot = p1, filename = paste0(opt$plot, drugSub, ".png"), 
       height = 5.25, width = 6.75)

```

#### Dose-response bafilomycin A1
```{r fig.height= 5, fig.width=8}
drugSub <- "Bafilomycin A1"

concInter <- ScreenB %>%
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  dplyr::select(conc) %>% unique() %>% unlist() %>% 
  as.vector() %>% round(3)
concInter[1:2] <- round(concInter[1:2], 1)
concInter[3:6] <- round(concInter[3:6], 2)
concInter[7:9] <- round(concInter[7:9], 3)

p1 <- ScreenB %>% 
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  ggplot(aes(x = conc, y = viab, group = sampleID)) +
  geom_line(size = 0.25) + 
  geom_point(size = 3.75, shape = 21, fill = "#64B5F6") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(limits = c(0, 1.2)) +
  scale_x_log10(labels = concInter, 
                breaks = concInter) +
  xlab("Concentration (µM)") + ylab("Viability") + 
  theme_classic() + ggtitle(paste("Screen B", drugSub, sep = " - ")) +
  theme(text = element_text(size = 17.5), 
        legend.key = element_blank(),
        plot.title = element_text(hjust=0.5),
        # axis.text.x = element_text(hjust = 1, angle = 45),
        axis.text = element_text(size = 17.5),
        axis.line = element_line(linewidth = 0.75),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
p1


ggsave(plot = p1, filename = paste0(opt$plot, drugSub, ".png"), 
       height = 5.25, width = 6.75)
```

#### Dose-response selinexor
```{r fig.height= 5, fig.width=7}
ScreenD <- drugList[["ScreenD"]] ## get ScreenD data

drugSub <- c("Selinexor")

concInter <- ScreenD %>%
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  dplyr::select(conc) %>% unique() %>% unlist() %>% as.vector()

p1 <- ScreenD %>% 
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  ggplot(aes(x = conc, y = viab, group = sampleID)) +
  geom_line(size = 0.25) + 
  geom_point(size = 3.75, shape = 21, fill = "#64B5F6") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(limits = c(0, 1.2)) +
  scale_x_log10(labels = concInter, 
                breaks = concInter) +
  xlab("Concentration (µM)") + ylab("Viability") + 
  theme_classic() + ggtitle(paste("Screen D", drugSub, sep = " - ")) +
  theme(text = element_text(size = 17.5), 
        legend.key = element_blank(),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 17.5),
        axis.line = element_line(linewidth = 0.75),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA))
p1

ggsave(plot = p1, filename = paste0(opt$plot, drugSub, ".png"), 
       height = 5, width = 6)
```

#### Dose-response venetoclax
```{r fig.height= 5, fig.width=7}
ScreenD <- drugList[["ScreenD"]] ## get ScreenD data

drugSub <- c("Venetoclax")

concInter <- ScreenD %>%
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  dplyr::select(conc) %>% unique() %>% unlist() %>% as.vector()

p1 <- ScreenD %>% 
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  ggplot(aes(x = conc, y = viab, group = sampleID)) +
  geom_line(size = 0.25) + 
  geom_point(size = 3.75, shape = 21, fill = "#64B5F6") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(limits = c(0, 1.2)) +
  scale_x_log10(labels = concInter, 
                breaks = concInter) +
  xlab("Concentration (µM)") + ylab("Viability") + 
  theme_classic() + ggtitle(paste("Screen D", drugSub, sep = " - ")) +
  theme(text = element_text(size = 17.5), 
        legend.key = element_blank(),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 17.5),
        axis.line = element_line(linewidth = 0.75),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA))
p1

ggsave(plot = p1, filename = paste0(opt$plot, drugSub, ".png"), 
       height = 5, width = 6)
```

#### Report average effect of key compounds - ScreenC
```{r}
## Report the average toxicity 
drugSub <- c("Fludarabine", "Nutlin-3a")

viabTab <- ScreenC %>% dplyr::filter(!name %in% c("PBS", "DMSO")) %>%
  group_by(diagnosis, name) %>%
  mutate(value = mean(viab.auc, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(dplyr::select(drugAnno, Drug, Target, Pathway), 
            by = c("name" = "Drug")) %>%
  dplyr::select(diagnosis, name, Target, Pathway, value) %>% unique()

viabTab %>% dplyr::filter(name %in% drugSub, diagnosis == "T-PLL")
```

#### Report average effect of key compounds - ScreenB
```{r}
## Report the average toxicity 
drugSub <- c("Ruxolitinib", "Duvelisib", "Fludarabine", "Mitoxantrone", "Cytarabine", "Panobinostat", "JIB04", "UNC0638", "OTX015", "JQ1", "Dinaciclib", "SNS-032", "Flavopiridol", "Selinexor", "Verdinexor", "Bafilomycin A1", "Thapsigargin", "Birinapant", "GDC-0152")

viabTab <- ScreenB %>% dplyr::filter(!name %in% c("PBS", "DMSO")) %>%
  group_by(diagnosis, name) %>%
  mutate(value = mean(viab.auc, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(dplyr::select(drugAnno, Drug, Target, Pathway), 
            by = c("name" = "Drug")) %>%
  dplyr::select(diagnosis, name, Target, Pathway, value) %>% unique()

viabTab %>% dplyr::filter(name %in% drugSub, diagnosis == "T-PLL") %>%
  arrange(Pathway, value)
```


#### Report average effect of key compounds - ScreenD
```{r}
## Report the average toxicity 
drugSub <- c("Fludarabine", "Cytarabine")

viabTab <- ScreenD %>% dplyr::filter(!name %in% c("PBS", "DMSO")) %>%
  group_by(diagnosis, name) %>%
  mutate(value = mean(viab.auc, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(dplyr::select(drugAnno, Drug, Target, Pathway), 
            by = c("name" = "Drug")) %>%
  dplyr::select(diagnosis, name, Target, Pathway, value) %>% unique()

viabTab %>% dplyr::filter(name %in% drugSub, diagnosis == "T-PLL")
```

#### Overview plot drug classes
```{r}
pathSub <- c("PI3K-AKT-mTOR", "Epigenome", "Cell Cycle", "Metabolism", "DNA Damage",
             "BCR/TCR", "MAPK", "ROS", "JAK-STAT", "Prg. Cell Death", "BRAF",
             "EGFR", "MEK", "Nuclear Traffic")

drugSub <- c("Venetoclax", "Navitoclax", "Birinapant", "Nutlin-3a", "Selinexor", "Verdinexor", "Navitoclax", "Fludarabine", 
               "Dacinostat", "Vorinostat", "Panobinostat", "Cytarabine", "Mitoxantrone", 
               "Dinaciclib", "SNS-032", "Ibrutinib", "Ruxolitinib", "Duvelisib", "UNC0638", 
               "JQ1", "OTX015", "JIB04", "Flavopiridol", "Bafilomycin A1", "Thapsigargin")

pList <- lapply(c("ScreenA", "ScreenB", "ScreenC", "ScreenD"), function(x) {
  Screen <- drugList[[x]] # get ScreenA data

  Screen$screen <- gsub(x = x, pattern = "Screen", replacement = "Screen ")
  ## compute the mean druf effect per patient x drug
  viabTab <- Screen %>% dplyr::filter(!is.na(name), !is.na(diagnosis)) %>%
    filter(diagnosis == "T-PLL") %>%
    dplyr::filter(!name %in% c("PBS", "DMSO")) %>%
    group_by(name, diagnosis) %>%
    mutate(value = mean(viab.auc, na.rm = TRUE)) %>%
    ungroup() %>%
    left_join(dplyr::select(drugAnno.red, Drug, Target, Pathway), 
              by = c("name" = "Drug")) %>%
    select(name, value, Pathway, screen) %>% unique() %>%
    mutate(Pathway = ifelse(name %in% c("Selinexor", "Verdinexor"), "Nuclear Traffic", Pathway)) %>%
    filter(Pathway %in% pathSub)
  

  viabTab %>%
    ggplot(aes(x = Pathway, y = value)) +
    #geom_boxplot() +
    geom_beeswarm(shape = 21, size = 2, stroke = 0.25, aes(fill = Pathway), alpha = 0.6) +
    geom_hline(yintercept = 1, linetype = "dashed") +
    ylim(0, 1.2) + xlab("") + ylab("Viability") +
    lgd +
    facet_wrap(. ~ screen, scales = "free_x") +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    geom_label_repel(data = viabTab[viabTab$name %in% drugSub, ], 
                     aes(label = name), 
                     max.overlaps = getOption("ggrepel.max.overlaps", default = 90),
                     segment.linetype = 2, force = 10, show.legend = FALSE, alpha = 0.8,
                     nudge_x = 0.15, nudge_y = -0.15) +
    theme(legend.position = "none") +
    scale_fill_manual(values = c("PI3K-AKT-mTOR" = "#1E88E5", "Epigenome" = "#009688", "Cell Cycle" = "#FF8F00", 
                                 "Metabolism" = "#E53935", "DNA Damage" = "#9575CD", "Nuclear Traffic" = "#1A237E",
                                 "BCR/TCR" = "#4FC3F7", "MAPK" = "#81C784", "ROS" = "#F06292", "JAK-STAT" = "#EF9A9A",
                                 "Prg. Cell Death" = "#C5E1A5", "BRAF" = "#B3E5FC", "EGFR" = "#FDD835", "MEK" = "#FFF9C4"))
  #viabTab %>% filter(name == "Selinexor")
})
pList
grid.arrange(grobs = pList)


```

#### Output session info
```{r pressure, echo=FALSE}
sessionInfo()
```

