---
title: "Figure 1"
author: "M. Pohly, J. Lu"
date: "`r Sys.Date()`"
output: html_document
---

#### Load libraries
```{r warning=FALSE, message=FALSE}
library(readxl)
library(tidyverse)
library(DrugScreenExplorer)
library(RColorBrewer)
library(ComplexHeatmap)
library(ggvenn)
library(ggtext)
library(parallel)
library(ggbeeswarm)
library(gridExtra)
library(circlize)
library(ggrepel)
```

#### Define variables
```{r}
opt <- list()
opt$drugscreen <- "data/submission/drugScreens_pseudo.RDS"
opt$druganno <- "misc/drugList_suppl.xlsx"
opt$plot <- "plots/Fig1/"

## Set theme
lgd <-  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5), 
        text = element_text(size = 15),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA), 
        legend.background = element_rect(fill='transparent',colour = NA), 
        strip.background = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
```

#### Load drug screen data and annotations
```{r}
drugList <- readRDS(opt$drugscreen)
drugAnno <- read_excel(opt$druganno) %>%
  dplyr::select(-Supplier) %>% unique()
```

##### Piechart - Screen A
```{r fig.height= 5, fig.width=5}
## Make annotation more uniform
drugAnno.red <- drugAnno %>%
  mutate(Pathway = ifelse(Pathway %in% c("HDAC", "HDM", "BET", "HMT", "HAT", "HMT, MYC", "NFKB, HDAC, DNA Damage"), "Epigenome", Pathway)) %>%
  mutate(Pathway = ifelse(Pathway %in% c("Apoptosis", "Necroptosis", "Autophagy"), "Prg. Cell Death", Pathway)) %>%
  mutate(Pathway = ifelse(Pathway %in% c("DNA Damage", "DNA Damage, NFKB", "Chemo"), "DNA Damage", Pathway)) %>%
  mutate(Pathway = ifelse(Pathway %in% c("EGFR", "Prg. Cell Death", "BRAF", "Chemo", "Epigenome", "JAK-STAT", "DNA Damage",
                                         "Cell Cycle", "BCR/TCR", "MEK", "PI3K-AKT-mTOR", "MAPK", "Metabolism", "ROS"), Pathway, "Other")) %>%
  group_by(Screen) %>% mutate(TotalDrugs = length(Drug)) %>% ungroup()

## ScreenA
pathOrder <- c("PI3K-AKT-mTOR", "Epigenome", "Cell Cycle", "Metabolism", "DNA Damage",
               "BCR/TCR", "MAPK", "ROS", "JAK-STAT", "Prg. Cell Death", "BRAF",
               "EGFR", "MEK", "Other")

colTab <- c("#1E88E5", "#009688", "#FF8F00", "#E53935", "#9575CD", "#4FC3F7", #"#4FC3F7",
                     "#81C784", "#F06292", "#EF9A9A", "#C5E1A5", "#B3E5FC", "#FDD835" , "#FFF9C4", "#1A237E")

p <- drugAnno.red %>% filter(Screen == "Screen A") %>%
  dplyr::count(Pathway, TotalDrugs) %>%
  mutate(percentage = n / TotalDrugs) %>%
#  arrange(desc(n)) %>%
  mutate(Pathway = factor(Pathway, levels = pathOrder)) %>%
  mutate(PathwayPer = paste(Pathway, " (", n, ")", sep = "")) %>%
  arrange(Pathway) %>%
  mutate(Pathway = factor(PathwayPer, levels = PathwayPer)) %>%
  ggplot(aes(x = "", y = percentage, fill = Pathway)) +
  geom_bar(stat = "identity", colour = "black") +
  #ggtitle("Drug Library") +
  ggtitle("") +
  coord_polar("y", start=0, direction = -1) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size = 12.5, face = "bold"),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        legend.background = element_rect(fill='transparent',colour = NA),
        strip.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  scale_fill_manual(values = colTab)
p

ggsave(plot = p, filename = paste0(opt$plot, "ScreenA_piechar.png"), height = 5,
       width = 5)
```

##### Piechart - Screen B
```{r fig.height= 5, fig.width=5}
## ScreenB
p <- drugAnno.red %>% filter(Screen == "Screen B") %>%
  dplyr::count(Pathway, TotalDrugs) %>%
  mutate(percentage = n / TotalDrugs) %>%
  #  arrange(desc(n)) %>%
  mutate(Pathway = factor(Pathway, levels = pathOrder)) %>%
  mutate(PathwayPer = paste(Pathway, " (", n, ")", sep = "")) %>%
  arrange(Pathway) %>%
  mutate(Pathway = factor(PathwayPer, levels = PathwayPer)) %>%
  ggplot(aes(x = "", y = percentage, fill = Pathway)) +
  geom_bar(stat = "identity", colour = "black") +
  #ggtitle("Drug Library") +
  ggtitle("") +
  coord_polar("y", start=0, direction = -1) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size = 12.5, face = "bold"),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        legend.background = element_rect(fill='transparent',colour = NA),
        strip.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  scale_fill_manual(values = colTab)
p

ggsave(plot = p, filename = paste0(opt$plot, "ScreenB_piechar.png"), height = 5,
       width = 5)
```

##### Piechart - Screen C
```{r fig.height= 5, fig.width=5}
## ScreenC
p <- drugAnno.red %>% filter(Screen == "Screen C") %>%
  dplyr::count(Pathway, TotalDrugs) %>%
  mutate(percentage = n / TotalDrugs) %>%
  #  arrange(desc(n)) %>%
  mutate(Pathway = factor(Pathway, levels = pathOrder)) %>%
  mutate(PathwayPer = paste(Pathway, " (", n, ")", sep = "")) %>%
  arrange(Pathway) %>%
  mutate(Pathway = factor(PathwayPer, levels = PathwayPer)) %>%
  ggplot(aes(x = "", y = percentage, fill = Pathway)) +
  geom_bar(stat = "identity", colour = "black") +
  #ggtitle("Drug Library") +
  ggtitle("") +
  coord_polar("y", start=0, direction = -1) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size = 12.5, face = "bold"),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        legend.background = element_rect(fill='transparent',colour = NA),
        strip.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  scale_fill_manual(values = colTab)
p

ggsave(plot = p, filename = paste0(opt$plot, "ScreenC_piechar.png"), height = 5,
         width = 5)
```

##### Piechart - Screen D
```{r fig.height= 5, fig.width=5}
## ScreenD
colTab <- c("#1E88E5", "#009688", "#FF8F00", "#9575CD", "#4FC3F7", #"#4FC3F7",
                     "#81C784", "#F06292", "#EF9A9A", "#C5E1A5", "#B3E5FC", "#FDD835" , "#FFF9C4", "#1A237E")

p <- drugAnno.red %>% filter(Screen == "Screen D") %>%
  dplyr::count(Pathway, TotalDrugs) %>%
  mutate(percentage = n / TotalDrugs) %>%
  #  arrange(desc(n)) %>%
  mutate(Pathway = factor(Pathway, levels = pathOrder)) %>%
  mutate(PathwayPer = paste(Pathway, " (", n, ")", sep = "")) %>%
  arrange(Pathway) %>%
  mutate(Pathway = factor(PathwayPer, levels = PathwayPer)) %>%
  ggplot(aes(x = "", y = percentage, fill = Pathway)) +
  geom_bar(stat = "identity", colour = "black") +
  #ggtitle("Drug Library") +
  ggtitle("") +
  coord_polar("y", start=0, direction = -1) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size = 12.5, face = "bold"),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        legend.background = element_rect(fill='transparent',colour = NA),
        strip.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  scale_fill_manual(values = colTab)
p

ggsave(plot = p, filename = paste0(opt$plot, "ScreenD_piechar.png"), height = 5,
       width = 5)

```

##### Piechart - Validation screen
```{r fig.height= 5.5, fig.width=5.5}
## Validation
colTab <- c("#1E88E5", "#009688", "#FF8F00", "#9575CD", "#4FC3F7", #"#4FC3F7",
                     "#81C784", "#EF9A9A", "#C5E1A5", "#FDD835" , "#FFF9C4", "#1A237E")
                     
p <- drugAnno.red %>% filter(Screen == "Screen E") %>%
  dplyr::count(Pathway, TotalDrugs) %>%
  mutate(percentage = n / TotalDrugs) %>%
  #  arrange(desc(n)) %>%
  mutate(Pathway = factor(Pathway, levels = pathOrder)) %>%
  mutate(PathwayPer = paste(Pathway, " (", n, ")", sep = "")) %>% 
  arrange(Pathway) %>%
  mutate(Pathway = factor(PathwayPer, levels = PathwayPer)) %>%
  ggplot(aes(x = "", y = percentage, fill = Pathway)) +
  geom_bar(stat = "identity", colour = "black") +
  #ggtitle("Drug Library") +
  ggtitle("") +
  coord_polar("y", start=0, direction = -1) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5), 
        text = element_text(size = 12.5, face = "bold"),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA), 
        legend.background = element_rect(fill='transparent',colour = NA), 
        strip.background = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  scale_fill_manual(values = colTab)
p

ggsave(plot = p, filename = paste0(opt$plot, "validation_piechar.png"), height = 5, 
       width = 5)
```

#### Fig 1C
##### Ranked toxicity - Screen A
```{r message=FALSE, fig.height= 6, fig.width=15}
ScreenA <- drugList[["ScreenA"]] # get ScreenA data

## index concentrations
ScreenA.ind <- mclapply(unique(ScreenA$name), mc.cores = 4, function(x) {
  message("Running ", x)
  ScreenA.x <- dplyr::filter(ScreenA, name == x) %>% 
    mutate(conc = as.numeric(conc))
  subTab <- ScreenA.x %>%
    dplyr::select(name, conc, concIndex) %>% arrange(desc(conc)) %>% unique()
 subTab$Index <- seq(1:nrow(subTab))
  message("Adding ", nrow(subTab), " indices")
  subTab %>% left_join(ScreenA.x, by = c("name", "conc", "concIndex"))
}) %>% bind_rows()

## compute the mean druf effect per patient x drug
viabTab <- ScreenA.ind %>% dplyr::filter(!is.na(name), !is.na(diagnosis)) %>%
  dplyr::filter(!name %in% c("PBS", "DMSO")) %>%
  group_by(patientID, name, diagnosis) %>%
  mutate(value = mean(viab.auc, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(dplyr::select(drugAnno, Drug, Target, Pathway), 
            by = c("name" = "Drug"))

topN <- viabTab$name %>% unique() %>% length() # define the number of drugs to plot
colorCode <- rev(brewer.pal(6,"Blues"))[1:6] # define colour scheme

## Subset to T-PLL patients
viabTab <- dplyr::filter(viabTab, diagnosis == "T-PLL") %>%
  mutate(Index = factor(Index)) %>%
  dplyr::select(patientID, diagnosis, name, Target, Pathway, viab.auc, value, viab, concIndex, Index) %>%
  unique()

## reshuffle before plotting
set.seed(100)
viabTab <- viabTab[sample(1:nrow(viabTab),nrow(viabTab)), ]

## Make ranked toxicity plot
p <- lapply(unique(viabTab$diagnosis), function(n) {
  eachTab <- dplyr::filter(viabTab, diagnosis == n)
  drugOrder <- group_by(eachTab, name) %>%
    summarise(medVal = median(value, na.rm = TRUE)) %>%
    arrange(medVal)
  drugOrder <- drugOrder[c(seq(1,topN)), ]
  eachTab <- dplyr::filter(eachTab, name %in% drugOrder$name) %>%
    mutate(name = factor(name, levels = drugOrder$name), 
           concIndex = as.factor(concIndex)) %>%
    group_by(name) %>% mutate(medVal = median(value, na.rm = TRUE)) %>%
    ungroup()
  drugOrder$rank <- seq(1:nrow(drugOrder))
  eachTab <- left_join(eachTab, dplyr::select(drugOrder, name, rank), by = "name")
  eachTab <- eachTab %>%
    group_by(name#, concIndex
             ) %>%
    mutate(viab.auc = mean(viab, na.rm = TRUE))
  p <- ggplot(eachTab, aes(x=rank,y = viab.auc#, fill = Index
                           )) +
    geom_jitter(alpha=0.8, shape = 21, size = 3, height = 0, fill = "#64B5F6") +
    #scale_color_manual(values = colorCode) + 
    scale_x_continuous(expand = c(0.005, 0), labels = c(100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200, 1300, 1400, 1500, 1600, 1700, 1800, 1900, 2000, 2100, 2200, 2300, 2400, 2500), 
                       breaks = c(100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200, 1300, 1400, 1500, 1600, 1700, 1800, 1900, 2000, 2100, 2200, 2300, 2400, 2500)) +
    ylab("Viability") + ylim(c(0,1.2)) + 
    xlab("Rank") + #ggtitle("ScreenA - Ranked Efficacy") +
    theme_classic() +
    geom_hline(yintercept = 1, linetype = "dashed") +
    theme(text = element_text(size = 17.5), #axis.text.x=element_blank(),
      #axis.ticks.x=element_blank(), 
      #axis.ticks = element_blank(), 
      legend.key = element_blank(),
      axis.line = element_line(linewidth = 0.75),
      #legend.position = "none",
      plot.title = element_text(hjust=0.5), 
      # plot.margin = unit(c(0.5, 0.5, 0.0, 0), "cm"), 
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      #legend.direction="horizontal",
      #legend.position = c(0.2, 0.25),
      panel.background = element_rect(fill = "transparent",colour = NA),
      plot.background = element_rect(fill = "transparent",colour = NA), 
      legend.background = element_rect(fill='transparent')) +
    guides(colour = guide_legend(override.aes = list(size=3.5), title="Concentration\nIndex"))
}) 
p

ggsave(plot = p[[1]], filename = paste0(opt$plot, "/ScreenA_overview_low.png"),
       height = 6.5, width = 20)
```

#### BH3 mimetics
```{r fig.height= 6.25, fig.width=7.25}
ScreenB <- drugList[["ScreenB"]] # get ScreenB data

viabTab <- ScreenB %>% dplyr::filter(!name %in% c("PBS", "DMSO")) %>%
  group_by(patientID, name, diagnosis) %>%
  mutate(value = mean(viab.auc, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(dplyr::select(drugAnno, Drug, Target, Pathway), 
            by = c("name" = "Drug")) %>%
  dplyr::select(patientID, sampleID, diagnosis, name, Target, Pathway, concIndex, 
         viab, value, viab.auc) %>% unique()

## Subset to BH3-mimetics in T-PLL 
bh3 <- c("Venetoclax", "Navitoclax", "Obatoclax mesylate", "Sabutoclax", "TW-37")
testTab <- viabTab %>% 
  dplyr::filter(diagnosis == "T-PLL", name %in% bh3) %>%
  dplyr::select(patientID, name, viab.auc) %>% unique() %>%
  mutate(name = factor(name, levels = bh3))

fit <- aov(viab.auc ~ name, testTab)
summary(fit)

## Compute p-value
lapply(bh3, function(x) {
  if(x == "Venetoclax") {
    fit.sc <- multcomp::glht(fit, linfct = multcomp::mcp(name = c(4, -1, -1, -1, -1))) # is venetoclax different from the rest?
    }
  if(x == "Navitoclax") {
    fit.sc <- multcomp::glht(fit, linfct = multcomp::mcp(name = c(4, -4, 0, 0, 0)))
    }
  if(x == "Obatoclax mesylate") {
    fit.sc <- multcomp::glht(fit, linfct = multcomp::mcp(name = c(4, 0, -4, 0, 0)))
    }
  if(x == "Sabutoclax") {
    fit.sc <- multcomp::glht(fit, linfct = multcomp::mcp(name = c(4, 0, 0, -4, 0)))
    }
  if(x == "TW-37") {
    fit.sc <- multcomp::glht(fit, linfct = multcomp::mcp(name = c(4, 0, 0, 0, -4)))
    }
  p.adj <- summary(fit.sc, test = multcomp::adjusted("BH"))$test$pvalues
  df <- data.frame(name = x, p.adj = p.adj)
}) %>% bind_rows()

## Get bcl2 affinity from references
bcl2bind <- data.frame(name = c("Venetoclax", "Navitoclax", "Obatoclax mesylate", 
                                "Sabutoclax", "TW-37"), 
                       bcl2aff = c(0.01, 1, 220, 310, 290)) # nM ScreenC
testTab <- testTab %>% left_join(bcl2bind, by = "name")

# fit <- lm(viab.auc ~ bcl2aff, testTab)
# summary(fit)

pVen <- viabTab %>%
  dplyr::filter(name %in% c("Venetoclax", "Navitoclax", "Obatoclax mesylate", 
                     "Sabutoclax", "TW-37"), diagnosis == "T-PLL") %>%
  left_join(bcl2bind, by = "name") %>%
  mutate(name = str_replace(name, "Obatoclax mesylate", "Obatoclax")) %>%
  mutate(name = factor(name, levels = c("Venetoclax", "Navitoclax", "Obatoclax", 
                "Sabutoclax", "TW-37"))) %>%
  dplyr::select(name, patientID, value, bcl2aff) %>% unique() %>%
  ggplot(aes(x = name, y = value, fill = log(bcl2aff))) +
  geom_boxplot(alpha = 0.75) +
  geom_beeswarm(size = 3.5, shape = 21, cex = 3, alpha = 0.75) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  xlab("") + ylab("Mean Viability") +
  ggtitle("Screen B - Activity of BH3-Mimetics") +
  theme_classic() +
  theme(text = element_text(size = 17.5),
    plot.title = element_text(hjust=0.5, size = 25), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 20),
    axis.text.y = element_text(size = 20),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
    axis.title = element_text(size = 22.5),
    legend.position = c(0.875, 0.25),
    axis.line = element_line(linewidth = 1),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA), 
        legend.background = element_rect(fill='transparent')) +
  scale_fill_continuous("Log2(Ki Bcl-2)")
pVen

ggsave(plot = pVen, filename = paste0(opt$plot, "diffVen.png"), 
       height = 6.5, width = 7.25)

```

### Dose-response fludarabine
```{r fig.height= 5, fig.width=6.5}
ScreenC <- drugList[["ScreenC"]] ## get ScreenC data

drugSub <- c("Fludarabine")

concInter <- ScreenC %>%
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  dplyr::select(conc) %>% unique() %>% unlist() %>% as.vector()

p1 <- ScreenC %>% 
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  ggplot(aes(x = conc, y = viab, group = sampleID)) +
  geom_line(size = 0.25) + 
  geom_point(size = 3.75, shape = 21, fill = "#64B5F6") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(limits = c(0, 1.2)) +
  scale_x_log10(labels = concInter, 
                breaks = concInter) +
  xlab("Concentration (µM)") + ylab("Viability") + 
  theme_classic() + ggtitle(paste("Screen C", drugSub, sep = " - ")) +
  theme(text = element_text(size = 17.5), 
        legend.key = element_blank(),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 17.5),
        axis.line = element_line(linewidth = 0.75),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA))
p1

ggsave(plot = p1, filename = paste0(opt$plot, drugSub, ".png"), 
       height = 5, width = 6)
```

#### Dose-response nutlin-3a
```{r fig.height= 5, fig.width=6.5}
ScreenC <- drugList[["ScreenC"]] ## get ScreenC data

drugSub <- c("Nutlin-3a")

concInter <- ScreenC %>%
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  dplyr::select(conc) %>% unique() %>% unlist() %>% as.vector()

p1 <- ScreenC %>% 
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  ggplot(aes(x = conc, y = viab, group = sampleID)) +
  geom_line(size = 0.25) + 
  geom_point(size = 3.75, shape = 21, fill = "#64B5F6") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(limits = c(0, 1.2)) +
  scale_x_log10(labels = concInter, 
                breaks = concInter) +
  xlab("Concentration (µM)") + ylab("Viability") + 
  theme_classic() + ggtitle(paste("Screen C", drugSub, sep = " - ")) +
  theme(text = element_text(size = 17.5), 
        legend.key = element_blank(),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 17.5),
        axis.line = element_line(linewidth = 0.75),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA))
p1

ggsave(plot = p1, filename = paste0(opt$plot, drugSub, ".png"), 
       height = 5, width = 6)

```

#### Dose-response birinapant
```{r fig.height= 5, fig.width=6.5}
drugSub <- "Birinapant"

concInter <- ScreenB %>%
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  dplyr::select(conc) %>% unique() %>% unlist() %>% 
  as.vector()
concInter[1:2] <- round(concInter[1:2], 1)
concInter[3:6] <- round(concInter[3:6], 2)
concInter[7:9] <- round(concInter[7:9], 3)

p1 <- ScreenB %>% 
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  ggplot(aes(x = conc, y = viab, group = sampleID)) +
  geom_line(size = 0.25) + 
  geom_point(size = 3.75, shape = 21, fill = "#64B5F6") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(limits = c(0, 1.2)) +
  scale_x_log10(labels = concInter, 
                breaks = concInter) +
  xlab("Concentration (µM)") + ylab("Viability") + 
  theme_classic() + ggtitle(paste("Screen B", drugSub, sep = " - ")) +
  theme(text = element_text(size = 17.5), 
        legend.key = element_blank(),
        plot.title = element_text(hjust=0.5),
        # axis.text.x = element_text(hjust = 1, angle = 45),
        axis.text = element_text(size = 17.5),
        axis.line = element_line(linewidth = 0.75),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA), 
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
p1

ggsave(plot = p1, filename = paste0(opt$plot, drugSub, ".png"), 
       height = 5.25, width = 6.75)

```

#### Dose-response bafilomycin A1
```{r fig.height= 5, fig.width=6.5}
drugSub <- "Bafilomycin A1"

concInter <- ScreenB %>%
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  dplyr::select(conc) %>% unique() %>% unlist() %>% 
  as.vector() %>% round(3)
concInter[1:2] <- round(concInter[1:2], 1)
concInter[3:6] <- round(concInter[3:6], 2)
concInter[7:9] <- round(concInter[7:9], 3)

p1 <- ScreenB %>% 
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  ggplot(aes(x = conc, y = viab, group = sampleID)) +
  geom_line(size = 0.25) + 
  geom_point(size = 3.75, shape = 21, fill = "#64B5F6") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(limits = c(0, 1.2)) +
  scale_x_log10(labels = concInter, 
                breaks = concInter) +
  xlab("Concentration (µM)") + ylab("Viability") + 
  theme_classic() + ggtitle(paste("Screen B", drugSub, sep = " - ")) +
  theme(text = element_text(size = 17.5), 
        legend.key = element_blank(),
        plot.title = element_text(hjust=0.5),
        # axis.text.x = element_text(hjust = 1, angle = 45),
        axis.text = element_text(size = 17.5),
        axis.line = element_line(linewidth = 0.75),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
p1


ggsave(plot = p1, filename = paste0(opt$plot, drugSub, ".png"), 
       height = 5.25, width = 6.75)
```

#### Dose-response selinexor
```{r fig.height= 5, fig.width=6.5}
ScreenD <- drugList[["ScreenD"]] ## get ScreenD data

drugSub <- c("Selinexor")

concInter <- ScreenD %>%
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  dplyr::select(conc) %>% unique() %>% unlist() %>% as.vector()

p1 <- ScreenD %>% 
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  ggplot(aes(x = conc, y = viab, group = sampleID)) +
  geom_line(size = 0.25) + 
  geom_point(size = 3.75, shape = 21, fill = "#64B5F6") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(limits = c(0, 1.2)) +
  scale_x_log10(labels = concInter, 
                breaks = concInter) +
  xlab("Concentration (µM)") + ylab("Viability") + 
  theme_classic() + ggtitle(paste("Screen D", drugSub, sep = " - ")) +
  theme(text = element_text(size = 17.5), 
        legend.key = element_blank(),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 17.5),
        axis.line = element_line(linewidth = 0.75),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA))
p1

ggsave(plot = p1, filename = paste0(opt$plot, drugSub, ".png"), 
       height = 5, width = 6)
```

#### Dose-response venetoclax
```{r fig.height= 5, fig.width=6.5}
ScreenD <- drugList[["ScreenD"]] ## get ScreenD data

drugSub <- c("Venetoclax")

concInter <- ScreenD %>%
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  dplyr::select(conc) %>% unique() %>% unlist() %>% as.vector()

p1 <- ScreenD %>% 
  dplyr::filter(name %in% c(drugSub), 
         diagnosis == "T-PLL") %>%
  ggplot(aes(x = conc, y = viab, group = sampleID)) +
  geom_line(size = 0.25) + 
  geom_point(size = 3.75, shape = 21, fill = "#64B5F6") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(limits = c(0, 1.2)) +
  scale_x_log10(labels = concInter, 
                breaks = concInter) +
  xlab("Concentration (µM)") + ylab("Viability") + 
  theme_classic() + ggtitle(paste("Screen D", drugSub, sep = " - ")) +
  theme(text = element_text(size = 17.5), 
        legend.key = element_blank(),
        plot.title = element_text(hjust=0.5),
        axis.text = element_text(size = 17.5),
        axis.line = element_line(linewidth = 0.75),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA))
p1

ggsave(plot = p1, filename = paste0(opt$plot, drugSub, ".png"), 
       height = 5, width = 6)
```

#### Report average effect of key compounds - ScreenC
```{r}
## Report the average toxicity 
drugSub <- c("Fludarabine", "Nutlin-3a")

viabTab <- ScreenC %>% dplyr::filter(!name %in% c("PBS", "DMSO")) %>%
  group_by(diagnosis, name) %>%
  mutate(value = mean(viab.auc, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(dplyr::select(drugAnno, Drug, Target, Pathway), 
            by = c("name" = "Drug")) %>%
  dplyr::select(diagnosis, name, Target, Pathway, value) %>% unique()

viabTab %>% dplyr::filter(name %in% drugSub, diagnosis == "T-PLL")
```

#### Report average effect of key compounds - ScreenB
```{r}
## Report the average toxicity 
drugSub <- c("Ruxolitinib", "Duvelisib", "Fludarabine", "Mitoxantrone", "Cytarabine", "Panobinostat", "JIB04", "UNC0638", "OTX015", "JQ1", "Dinaciclib", "SNS-032", "Flavopiridol", "Selinexor", "Verdinexor", "Bafilomycin A1", "Thapsigargin", "Birinapant", "GDC-0152")

viabTab <- ScreenB %>% dplyr::filter(!name %in% c("PBS", "DMSO")) %>%
  group_by(diagnosis, name) %>%
  mutate(value = mean(viab.auc, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(dplyr::select(drugAnno, Drug, Target, Pathway), 
            by = c("name" = "Drug")) %>%
  dplyr::select(diagnosis, name, Target, Pathway, value) %>% unique()

viabTab %>% dplyr::filter(name %in% drugSub, diagnosis == "T-PLL") %>%
  arrange(Pathway, value)
```

#### Report average effect of key compounds - ScreenD
```{r}
## Report the average toxicity 
drugSub <- c("Fludarabine", "Cytarabine")

viabTab <- ScreenD %>% dplyr::filter(!name %in% c("PBS", "DMSO")) %>%
  group_by(diagnosis, name) %>%
  mutate(value = mean(viab.auc, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(dplyr::select(drugAnno, Drug, Target, Pathway), 
            by = c("name" = "Drug")) %>%
  dplyr::select(diagnosis, name, Target, Pathway, value) %>% unique()

viabTab %>% dplyr::filter(name %in% drugSub, diagnosis == "T-PLL")
```

#### Output session info
```{r pressure, echo=FALSE}
sessionInfo()
```

