---
title: "Figure 2"
author: "M. Pohly, J. Lu"
date: "`r Sys.Date()`"
output: html_document
---

#### Load libraries
```{r warning=FALSE, message=FALSE}
library(readxl)
library(tidyverse)
library(jyluMisc)
library(DrugScreenExplorer)
library(ggbeeswarm)
library(ggrepel)
library(ComplexHeatmap)
library(scales)
library(parallel)
library(gridExtra)
```

#### Define variables
```{r}
opt <- list()
opt$drugscreen <- "data/submission/drugScreens_pseudo.RDS"
opt$druganno <- "misc/drugList_suppl.xlsx"
opt$plot <- "plots/Fig2/"
```

#### Load drug screen data and annotations
```{r}
drugList <- readRDS(opt$drugscreen)
drugAnno <- read_excel(opt$druganno) %>%
  dplyr::select(-Supplier, -Screen) %>% unique()
```

#### Volcano plot T-PLL vs CLL - Screen A
```{r fig.height= 5.5, fig.width=5.5}
embl2014 <- drugList[["ScreenA"]]

## Make t-test function
tTest <- function(value, type) {
  res <- t.test(value ~ factor(type), equal.var = TRUE, na.action = na.exclude)
  tibble(p = res$p.value,
         diff = res$estimate[[2]] - res$estimate[[1]])
}

compDiag <- "CLL"
diagSelected <- c(compDiag, "T-PLL")

screenDat <- embl2014

pTab <- lapply(diagSelected[diagSelected != c(compDiag)], function(diag) {
  print(diag)
  testTab <- filter(screenDat, diagnosis %in% c(compDiag, diag),
                    !name %in% c("DMSO", "empty")) %>%
    #filter(name %in% c("Venetoclax")) %>%
    mutate(diagnosis = factor(diagnosis, levels = c(compDiag, diag)), 
           viab = viab.auc) %>%
    select(patientID, diagnosis, name, viab) %>% filter(!is.na(viab)) %>% unique()
  res <- group_by(testTab, name) %>% do(tTest(.$viab, .$diagnosis))
  res$diagnosis <- diag
  res
}) %>% bind_rows() %>% ungroup() %>% mutate(p.adj = p.adjust(p, method = "BH"))

plotTab <- pTab %>% mutate(p = ifelse(p < 1e-12, 1e-12, p)) %>%
  mutate(pSign = -log10(p)*sign(diff)) %>%
  ungroup()

plotTab$diffCol <- "ns"
plotTab[plotTab[, "p"] < 0.05 & plotTab[, "diff"] < -0.05, "diffCol"] <- "higher_sen"
plotTab[plotTab[, "p"] < 0.05 & plotTab[, "diff"] > 0.05, "diffCol"] <- "lower_sen"

drugSelec <- c("Nutlin-3a", "Abt-737", "Palbociclib", "Venetoclax", "Prednisolone", 
               "Thapsigargin", "Bortezomib", "Duvelisib", "Vindesine", "CCT241533", 
               "Ibrutinib", "Beclomethasone diproprionate", "Betametasone", "GSK1627798A", 
               "GW352430A", "Fluorometholone", "Salbutamol Sulfate",
               "Dexamethasone", #"SAM002589973", "GW440139B", "GSK1292139B", 
               "SB-437013", "GW680191X", "SB-601273",
               "GW461104B", "GW591947A",  "GW856805X",
               "Cytarabine", "GSK1307810A", "Serdemetan", #"Rolipram", 
               "GW780159X", "GSK1627798A", "Thiostrepton")

## Rename salbutamol
plotTab <- plotTab %>% mutate(name = str_replace(name, pattern = "SALBUTAMOL SULFATE", replacement = "Salbutamol Sulfate"))

p1 <- plotTab %>%
  ggplot(aes(x = diff, y = -log10(p))) +
  geom_point(shape = 21, size = 2.5, aes(fill = diffCol)) +
  #  geom_text(data = filter(plotTab, name %in% c(drugSelec)), aes(label = name), 
  #            nudge_y = 0.75, size = 2.75) +
  #geom_text(label = "p = 0.05", x = -0.45, y = 3.5, colour = "gray20") +
  geom_hline(yintercept = 1.3, linetype = "dashed") +
  geom_vline(xintercept = 0.05, linetype = "dashed") +
  geom_vline(xintercept = -0.05, linetype = "dashed") +
  scale_x_reverse(limits = c(0.5, -0.5)) +
  xlab("Viability Difference") + ylab("-Log10(p)") +
  ggtitle(paste0("Screen A - T-PLL vs ", compDiag)) +
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5, size = 17.5), 
        axis.title = element_text(size = 15),
        axis.text = element_text(size = 12.5),
        axis.line = element_line(linewidth = 0.75),
        legend.position = "none",
        legend.key = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        legend.background = element_rect(fill='transparent')) +
  scale_fill_manual(values = c("higher_sen" = "#1976D2", "lower_sen" = "#F44336", "ns" = "grey")) +
  geom_label_repel(data = filter(plotTab, name %in% c(drugSelec), p < 0.05, diff > 0.05), 
                   aes(x = diff, y = -log10(p), label = name), size = 3, alpha = 0.8, 
                   max.overlaps = getOption("ggrepel.max.overlaps", default = 30),
                   segment.linetype = 2, force = 10, show.legend = FALSE, xlim = c(-0.175, -0.5),
                   nudge_y = 0.5) + 
  geom_label_repel(data = filter(plotTab, name %in% c(drugSelec), p < 0.05, diff < - 0.05), 
                   aes(x = diff, y = -log10(p), label = name), size = 3, alpha = 0.8, 
                   max.overlaps = getOption("ggrepel.max.overlaps", default = 3),
                   segment.linetype = 2, force = 10, show.legend = FALSE, xlim = c(0.025, 0.5),
                   nudge_y = 0.5)

p1

ggsave(plot = p1, file = paste0(opt$plot, "vol_embl2014_", compDiag, ".png"), 
       height = 5.5, width = 5.5)

## Output p-value and diff for compounds mentioned in the text
plotTab %>% filter(name %in% c("Nutlin-3a"))
```


#### Volcano plot T-PLL vs CLL - Screen B
```{r fig.height= 5.5, fig.width=5.5}
embl2016 <- drugList[["ScreenB"]]

compDiag <- "CLL"
diagSelected <- c(compDiag, "T-PLL")

screenDat <- embl2016
pTab <- lapply(diagSelected[diagSelected != c(compDiag)], function(diag) {
  print(diag)
  testTab <- filter(screenDat, diagnosis %in% c(compDiag, diag),
                    !name %in% c("DMSO", "empty")) %>%
    group_by(patientID, name) %>%
    #filter(name %in% c("Venetoclax")) %>%
    mutate(diagnosis = factor(diagnosis, levels = c(compDiag, diag)), 
           viab = mean(normVal.sigm, na.rm = TRUE)) %>% ungroup() %>%
    select(patientID, diagnosis, name, viab) %>% filter(!is.na(viab)) %>% unique()
  res <- group_by(testTab, name) %>% do(tTest(.$viab, .$diagnosis))
  res$diagnosis <- diag
  res
}) %>% bind_rows() %>% ungroup() %>% mutate(p.adj = p.adjust(p, method = "BH"))

plotTab <- pTab %>% mutate(p = ifelse(p < 1e-8, 1e-8, p)) %>%
  mutate(pSign = -log10(p)*sign(diff)) %>%
  ungroup()

plotTab$diffCol <- "ns"
plotTab[plotTab[, "p"] < 0.05 & plotTab[, "diff"] < -0.05, "diffCol"] <- "higher_sen"
plotTab[plotTab[, "p"] < 0.05 & plotTab[, "diff"] > 0.05, "diffCol"] <- "lower_sen"

drugSelec <- c("QS11", "Verdinexor", "Selinexor","Thalidomide", "Birinapant", "Bafilomycin A1",
                                      "Thapsigargin", "Prednisolone", "Selumetinib", "Venetoclax", "Navitoclax", 
                                      "Fludarabine", "Pentostatin", "Ibrutinib", "Duvelisib", "Idelalisib", "GDC-0152",
                                      "Doxorubicin", #"ONO-4059", "CNX-774",  "PD0325901", 
                                      "Dexamethasone", "Motolimod", 
                                      "Dasatinib", "Swainsonine", "AZD9291", #"Lestaurtinib",# "AGI-5198", "EPZ-5676", 
                                      "AZD7762", "Iniparib")

p1 <- plotTab %>% 
  ggplot(aes(x = diff, y = -log10(p))) +
  geom_point(shape = 21, size = 2.5, aes(fill = diffCol)) +
  geom_hline(yintercept = 1.3, linetype = "dashed") +
  geom_vline(xintercept = 0.05, linetype = "dashed") +
  geom_vline(xintercept = -0.05, linetype = "dashed") +
  scale_x_reverse(limits = c(0.35, -0.35)) +
  xlab("Viability Difference") + ylab("-Log10(p)") +
  #ggtitle(paste0("EMBL2016 - Diff. Sen. ", compDiag)) +
  ggtitle(paste0("Screen B - T-PLL vs ", compDiag)) +
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5, size = 17.5), 
        axis.title = element_text(size = 15),
        axis.text = element_text(size = 12.5),
        axis.line = element_line(linewidth = 0.75),
        legend.position = "none",
        legend.key = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        legend.background = element_rect(fill='transparent')) +
  scale_fill_manual(values = c("higher_sen" = "#1976D2", "lower_sen" = "#F44336", "ns" = "grey")) +
  geom_label_repel(data = filter(plotTab, name %in% c(drugSelec), p < 0.05, diff > 0.05), 
                   aes(x = diff, y = -log10(p), label = name), size = 3, alpha = 0.8, 
                   max.overlaps = getOption("ggrepel.max.overlaps", default = 5),
                   segment.linetype = 2, force = 10, show.legend = FALSE, xlim = c(-0.1, -0.35),
                   nudge_y = 0.5) + 
  geom_label_repel(data = filter(plotTab, name %in% c(drugSelec), p < 0.05, diff < - 0.05), 
                   aes(x = diff, y = -log10(p), label = name), size = 3, alpha = 0.8, 
                   max.overlaps = getOption("ggrepel.max.overlaps", default = 5),
                   segment.linetype = 2, force = 10, show.legend = FALSE, xlim = c(0.125, 0.35),
                   nudge_y = 0.5)
p1

ggsave(plot = p1, file = paste0(opt$plot, "vol_embl2016_", compDiag, ".png"), 
         height = 5.5, width = 5.5)
## Output p-value and diff for compounds mentioned in the text
plotTab %>% filter(name %in% c("Pentostatin", "Venetoclax", "Navitoclax", "Ibrutinib", "Thalidomide", "Bafilomycin A1", "Birinapant", "GDC-0152"))

```

#### Volcano plot T-PLL vs CLL - Screen D
```{r fig.height= 5.5, fig.width=5.5}
cps1000 <- drugList[["ScreenD"]]

compDiag <- "CLL"
diagSelected <- c(compDiag, "T-PLL")

screenDat <- cps1000
pTab <- lapply(diagSelected[diagSelected != c(compDiag)], function(diag) {
  print(diag)
  testTab <- filter(screenDat, diagnosis %in% c(compDiag, diag),
                    !name %in% c("DMSO", "empty")) %>%
    mutate(diagnosis = factor(diagnosis, levels = c(compDiag, diag)), 
           viab = viab.auc) %>%
    select(patientID, diagnosis, name, viab) %>% filter(!is.na(viab)) %>% unique()
  res <- group_by(testTab, name) %>% do(tTest(.$viab, .$diagnosis))
  res$diagnosis <- diag
  res
}) %>% bind_rows() %>% ungroup() %>% mutate(p.adj = p.adjust(p, method = "BH"))

avg.viab <- screenDat %>% filter(diagnosis == "T-PLL") %>%
  group_by(name) %>%
  summarise(viab.auc = mean(viab, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(toxicity = 1 - viab.auc)

plotTab <- pTab %>% mutate(p = ifelse(p < 1e-12, 1e-12, p)) %>%
  mutate(pSign = -log10(p)*sign(diff)) %>%
  ungroup() %>%
  left_join(avg.viab, by = "name")

plotTab$diffCol <- "ns"
plotTab[plotTab[, "p"] < 0.05 & plotTab[, "diff"] < -0.05, "diffCol"] <- "higher_sen"
plotTab[plotTab[, "p"] < 0.05 & plotTab[, "diff"] > 0.05, "diffCol"] <- "lower_sen"

drugSelec <- c("Fludarabine", "Venetoclax", "Navitoclax", "Duvelisib", "Idelalisib", "Birinapant", 
               "Dasatinib", 
               "Ibrutinib", "QS11", "Bafilomycin A1", "Dexamethasone", "Predisolon", "Nutlin-3a", 
               "Selinexor", "RO5963", "Doxorubicine", "Palbociclib", "TW-37", "Rapamycin", 
               "ONO-4059", "PRT062607", "Ganetespib", "Onalespib", "Foretinib", "MI-503")

plotTab %>% filter(p < 0.05, diff < 0) %>% arrange(p)

if(compDiag == "CLL") {x_lim <- c(0.35, -0.35)}

p1 <- plotTab %>%
  ggplot(aes(x = diff, y = -log10(p))) +
  geom_point(shape = 21, size = 2.5, aes(fill = diffCol)) +
  geom_hline(yintercept = 1.3, linetype = "dashed") +
  geom_vline(xintercept = 0.05, linetype = "dashed") +
  geom_vline(xintercept = -0.05, linetype = "dashed") +
  scale_x_reverse(limits = x_lim) +
  xlab("Viability Difference") + ylab("-Log10(p)") +
  ggtitle(paste0("Screen D - T-PLL vs ", compDiag)) +
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5, size = 17.5), 
        axis.title = element_text(size = 15),
        axis.text = element_text(size = 12.5),
        axis.line = element_line(linewidth = 0.75),
        legend.position = "none",
        legend.key = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        legend.background = element_rect(fill='transparent')) +
  scale_fill_manual(values = c("higher_sen" = "#1976D2", "lower_sen" = "#F44336", "ns" = "grey")) +
  geom_label_repel(data = filter(plotTab, name %in% c(drugSelec), p < 0.05, diff > 0.05), 
                   aes(x = diff, y = -log10(p), label = name), size = 3, alpha = 0.8, 
                   max.overlaps = getOption("ggrepel.max.overlaps", default = 5),
                   segment.linetype = 2, force = 10, show.legend = FALSE, xlim = c(-0.125, -0.35),
                   nudge_y = 0.5) + 
  geom_label_repel(data = filter(plotTab, name %in% c(drugSelec), p < 0.05, diff < - 0.05), 
                   aes(x = diff, y = -log10(p), label = name), size = 3, alpha = 0.8, 
                   max.overlaps = getOption("ggrepel.max.overlaps", default = 5),
                   segment.linetype = 2, force = 10, show.legend = FALSE, xlim = c(0.125, 0.35),
                   nudge_y = 0.5)
p1

ggsave(plot = p1, file = paste0(opt$plot, "vol_cps1000_", compDiag, ".png"), 
       height = 5.5, width = 5.5)

## Output p-value and diff for compounds mentioned in the text
plotTab %>% filter(name %in% c("Fludarabine", "Selinexor"))
```

#### Heatmap comparison to CLL - Validation screen
```{r fig.height= 6, fig.width=11}
t_lymphoma <- drugList[["ScreenE"]]

## Drugs of interest
drugSub <- c("Bafilomycin A1", "Pomalidomide", "Birinapant", 
             "Selinexor", "Nutlin-3a", "Fludarabine", "Venetoclax", 
             "Navitoclax", "Tofacitinib", "Dacinostat", "Ibrutinib", "Everolimus", 
             "Ruxolitinib", "Duvelisib", "GDC-0152", "Birinapant", 
             "Doxorubicin")

screenDat <- t_lymphoma %>%
  filter(name %in% drugSub) %>%
  filter(!c(name == "Birinapant" & screen == "T_lymphoma_combi")) %>% # remove birinapant at 5 concentrations and only use the 10 concentration from the ECHO plates
  filter(!screen == "T_lymphoma_combi")

## Summarise the effect of technical replicates
screenDat <- screenDat %>%
  group_by(patientID, diagnosis, screen, name, concIndex) %>%
  summarise(viab = mean(viab, na.rm = TRUE), viab.auc = mean(viab.auc, na.rm = TRUE)) %>%
  ungroup()

## Rename the ibrutinib combination with compound 26 to save space in figures.
screenDat <- screenDat %>%
  separate(name, into = c("col1", "col2"), sep = "[|]") %>%
  mutate(col2 = str_replace(col2, pattern = "Compound 26 \\(0.1\\)", replacement = "Compound 26")) %>%
  mutate(name = ifelse(!is.na(col2), paste0(col1, "|", col2), col1)) %>%
  select(-col1, -col2)

compDiag <- "CLL"
diagSelected <- c(compDiag, "T-PLL", "T-LGL", "Sezary", "PTCL", "MCL")

pTab <- lapply(diagSelected[diagSelected != c(compDiag)], function(diag) {
  print(diag)
  testTab <- filter(screenDat, diagnosis %in% c(compDiag, diag)) %>%
    mutate(diagnosis = factor(diagnosis, levels = c(compDiag, diag)), 
           viab = viab)
  res <- group_by(testTab, name, concIndex) %>% do(tTest(.$viab, .$diagnosis))
  res$diagnosis <- diag
  res
}) %>% bind_rows() %>% ungroup() %>% mutate(p.adj = p.adjust(p, method = "BH"))

plotTab <- pTab %>% mutate(p = ifelse(p < 1e-05, 1e-05, p)) %>%
  mutate(pSign = -log10(p)*sign(diff), 
         sigSign = ifelse(p < 0.05, "*", "")) %>%
  ungroup()

plotTab <- filter(plotTab, name %in% drugSub, !concIndex %in% c(6:10)) # remove lowest 5 concentrations of birinapant and GDC-0152

pMat <- mutate(plotTab, diagConc = paste0(diagnosis,"_",concIndex)) %>%
  dplyr::select(name, pSign, diagConc) %>%
  spread(key = diagConc, value = pSign) %>% data.frame() %>%
  column_to_rownames("name") %>% as.matrix()

drugOrder <- rownames(pMat)[hclust(dist(pMat), method = "ward.D2")$order]
plotTab <- mutate(plotTab, name = factor(name, levels = rev(drugOrder)))

## Global overview
p1 <- plotTab %>% mutate(diagnosis = factor(diagnosis, levels = c("MCL", "T-PLL", "T-LGL", "PTCL", "Sezary"))) %>%
  ggplot(aes(x=factor(concIndex), y = name, fill = pSign)) + 
  geom_tile(size = 0.1, color = "black") + facet_wrap(~ diagnosis, nrow = 1) + 
  scale_fill_gradient2(high = "#F44336", mid = "white", low = "#1976D2", midpoint = 0,
                       name = "-Log10(p) \nwith Direction", limits = c(-4, 4), oob = squish) +
  theme_classic() + theme(strip.text = element_text(face = "bold"),
                          axis.text.y = element_text(size =12)) +
  xlab("Concentration Index") + #ylab("Drug name") + 
  ggtitle(paste("Drug Activity Compared to", compDiag,
                sep = " ")) +
  theme(text = element_text(size = 17.5), 
        plot.title = element_text(hjust = 0.5, size = 16),
        #axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 12.5),
        legend.title = element_text(size = 12.5),
        legend.text = element_text(size = 12),
        strip.text = element_text(size=12.5),
        strip.background = element_blank(),
        legend.key = element_blank(),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA), 
        legend.background = element_rect(fill='transparent'), 
        panel.border = element_rect(colour = "black", fill=NA, size=1))
p1

ggsave(plot = p1, filename = paste0(opt$plot, "tlymph_heatmap.png"), 
       height = 6, width = 11)

## Define drug combinations
drugComb <- c("Birinapant|Necrostatin-1 (25)", "Birinapant|QVD-Oph (25)", "Birinapant|Necrostatin-1 (25)|QVD-Oph (25)", 
              "Birinapant|Necrostatin-1 (12.5)", "Birinapant|QVD-Oph (12.5)", "Birinapant|Necrostatin-1 (12.5)|QVD-Oph (12.5)",
              "GDC-0152|Necrostatin-1 (25)", "GDC-0152|QVD-Oph (25)", "GDC-0152|Necrostatin-1 (25)|QVD-Oph (25)", "GDC-0152|Necrostatin-1 (12.5)", 
              "GDC-0152|QVD-Oph (12.5)", "GDC-0152|Necrostatin-1 (12.5)|QVD-Oph (12.5)", "Birinapant|Ipatasertib (2)", 
              "Birinapant|Ruxolitinib (2)", "Birinapant|Dacinostat (0.04)", "Birinapant|Venetoclax (0.04)",
              "Birinapant|Bafilomycin_A1 (0.08)", "Birinapant|NSA (0.8)", "Birinapant|NSA (2)", 
              "Birinapant|NSA (5)", "Birinapant|NSA (12.5)", "Birinapant|NSA (0.8)|QVD-Oph (25)", 
              "Birinapant|NSA (2)|QVD-Oph (25)", "Birinapant|NSA (5)|QVD-Oph (25)", "Birinapant|NSA (12.5)|QVD-Oph (25)", 
              "Birinapant + Ipatasertib 2µM", "Birinpant + Ruxolitinib 2µM", "Birinapant + Bafilomycin A1 0.08µM", 
              "Birinapant + NSA 0.8µM",
              "Birinapant + NSA 0.8µM + QVD-Oph 25µM", "Birinapant + NSA 2µM",  "Birinapant + NSA 5µM", 
              "Birinapant + NSA 2µM + QVD-Oph 25µM", "Birinapant + NSA 5μM + QVD-Oph 25μM", 
              "DMSO", "empty", "Necrostatin-1|QVD-Oph")

#### Reporting p-values for differential response
resTab <- lapply(diagSelected[diagSelected != c(compDiag)], function(diag) {
  print(diag)
  testTab <- filter(screenDat, diagnosis %in% c(compDiag, diag),
                    !name %in% c("DMSO", "empty", drugComb)) %>%
    #filter(name %in% drugSub) %>%
    mutate(diagnosis = factor(diagnosis, levels = c(compDiag, diag))) %>%
    select(name, viab.auc, diagnosis, patientID) %>% unique() 
  # testTab <- testTab %>% group_by(patientID, name) %>%
  #   mutate(viab.auc = mean(viab, na.rm = TRUE)) %>% ungroup() 
  res <- group_by(testTab, name) %>% do(tTest(.$viab.auc, .$diagnosis))
  res$diagnosis <- diag
  res
}) %>% bind_rows() %>% ungroup() %>% mutate(p.adj = p.adjust(p, method = "BH"))

## For T-PLL report raw p-values, since validation
resTab %>% filter(name %in% c("Nutlin-3a", "Birinapant", "Bafilomycin A1", "Selinexor", 
                     "Fludarabine", "Venetoclax", "Ruxolitinib", "Pomalidomide"), 
                  diagnosis %in% c("T-PLL"))

```

#### Boxplots - Screen E
```{r fig.height = 7.5, fig.width = 8}
t_lymphoma <- drugList[["ScreenE"]]

## Define drugs of interest to plot
drugSub <- c("Birinapant", "Selinexor", 
             "Bafilomycin A1", "Nutlin-3a")

## Plot average viability
plotTab <- t_lymphoma %>% 
  filter(name %in% drugSub, screen == "T_lymphoma", 
         !diagnosis %in% c("DLBCL", "MZL")) %>%
  group_by(patientID, name) %>%
  mutate(meanViab = mean(viab, na.rm = TRUE)) %>%
  ungroup() %>% select(name, diagnosis, patientID, meanViab) %>% 
  unique() %>% 
  mutate(diagnosis = factor(diagnosis, levels = c("CLL", "MCL", "T-PLL", "T-LGL", "Sezary", "PTCL", "AITL")), 
         name = factor(name, levels = drugSub))

plotTab.sen <- plotTab %>% mutate(meanViab = ifelse(meanViab > 1.25, 1.25, meanViab))

pT <- plotTab.sen %>%
  ggplot(aes(x = diagnosis, y = meanViab)) +
  geom_boxplot(aes(fill = diagnosis), alpha = 0.8) +
  geom_beeswarm(shape = 21, cex = 3, size = 2.25, aes(fill = diagnosis), 
                data = plotTab.sen[plotTab.sen$meanViab < 1.25, ]) +
  geom_beeswarm(shape = 24, cex = 3, size = 2.25, aes(fill = diagnosis), 
                data = plotTab.sen[plotTab.sen$meanViab >= 1.25, ]) +
  xlab("") + ylab("Avg. Viabibility") + #ylim(0, 1.5) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_bw() +
  theme(text = element_text(size = 17.5),
        plot.title = element_text(hjust = 0.5, size = 15), 
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA), 
        legend.background = element_rect(fill='transparent'), 
        strip.background = element_blank(), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.ticks.x = element_blank(),
        strip.text.x = element_text(size = 17.5), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        legend.title = element_blank(), 
        panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  facet_wrap(. ~ name, nrow = 2, scales = "free_x") +
  scale_fill_manual(values = c("MCL" = "#EF5350", "CLL" = "grey40","Other T-cell" = "#A5D6A7", "T-PLL" =  "#64B5F6", 
                               "T-LGL" = "#FFC107", "Sezary" = "#E57373", "PTCL" = "#C5E1A5", "AITL" = "#E1BEE7"))

pT

ggsave(plot = pT, paste0(opt$plot, "tlymph_comp.png"), 
       height = 8, width = 8)


# pT.long <- plotTab.sen %>%
#   filter(!name == "Nutlin-3a") %>%
#   ggplot(aes(x = diagnosis, y = meanViab)) +
#   geom_boxplot(aes(fill = diagnosis), alpha = 0.8) +
#   geom_beeswarm(shape = 21, cex = 3, size = 2.25, aes(fill = diagnosis),
#                 data = plotTab.sen[plotTab.sen$meanViab < 1.25 & !plotTab.sen$name %in% c("Nutlin-3a"), ]) +
#   geom_beeswarm(shape = 24, cex = 3, size = 2.25, aes(fill = diagnosis),
#                 data = plotTab.sen[plotTab.sen$meanViab >= 1.25 & !plotTab.sen$name %in% c("Nutlin-3a"), ]) +
#   xlab("") + ylab("Avg. Viabibility") + #ylim(0, 1.5) +
#   geom_hline(yintercept = 1, linetype = "dashed") +
#   theme_bw() +
#   theme(text = element_text(size = 22.5),
#         plot.title = element_text(hjust = 0.5, size = 15),
#         panel.background = element_rect(fill = "transparent",colour = NA),
#         plot.background = element_rect(fill = "transparent",colour = NA),
#         legend.background = element_rect(fill='transparent'),
#         strip.background = element_blank(),
#         axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
#         axis.ticks.x = element_blank(),
#         strip.text.x = element_text(size = 22.5),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         legend.title = element_blank(),
#         legend.position = "none",
#         panel.border = element_rect(colour = "black", fill=NA, size=1.5)) +
#   facet_wrap(. ~ name, nrow = 1, scales = "free_x") +
#   scale_fill_manual(values = c("MCL" = "#EF5350", "CLL" = "grey40","Other T-cell" = "#A5D6A7", "T-PLL" =  "#64B5F6",
#                                "T-LGL" = "#FFC107", "Sezary" = "#E57373", "PTCL" = "#C5E1A5", "AITL" = "#E1BEE7"))
# 
# pT.long
# 
# ggsave(plot = pT.long, paste0(opt$plot, "tlymph_comp_long.png"),
#        height = 5, width = 12)
```

#### Dose response - curve birinapant
```{r fig.height= 5, fig.width=6.5}
drugSelec <- "Birinapant"
screenSelec <- "T_lymphoma"
#screenSelec <- "T_lymphoma_combi"

## Compute average in case of multiple samples
t_lymphoma <- t_lymphoma %>%
  group_by(patientID, name, concentration, screen) %>%
  mutate(viab = mean(viab)) %>%
  select(patientID, diagnosis, name, screen, concentration, concIndex, viab) %>%
  unique() %>%
  ungroup()

concInter <- t_lymphoma %>% 
  filter(name %in% c(drugSelec), screen == screenSelec, 
         diagnosis == "T-PLL") %>%
  pull(concentration) %>% unique() %>% sort()

## Round low values for plotting
concInter[1] <- round(concInter[1], 5)
concInter[2:3] <- round(concInter[2:3], 4)

p3 <- t_lymphoma %>% 
  filter(name == drugSelec, diagnosis %in% c("T-PLL", "AITL", "PTCL", "T-LGL", "Sezary"),
         screen == screenSelec) %>%
  mutate(diagnosis = ifelse(diagnosis %in% c("T-PLL"), "T-PLL", "T-NHL")) %>%
  ggplot(aes(x = concentration, y = viab, group = patientID, colour = diagnosis, fill = diagnosis)) +
  geom_line(linewidth = 0.25, colour = "black") + 
  xlab("Concentration (µM)") + ylab("Viability") + 
  geom_point(shape = 21, size = 3.25, colour = "black") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(limits = c(0, 1.5)) +
  scale_x_log10(labels = concInter, 
                breaks = concInter) +
  theme_classic() + ggtitle(drugSelec) +
  theme(text = element_text(size = 17.5),
        legend.key = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        plot.title = element_text(hjust=0.5, size = 17.5), 
        legend.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.title = element_blank(),
        axis.line = element_line(linewidth = 0.75),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA)) +
  #scale_colour_manual(values = c("T-PLL" = "#CE93D8", "Other T-cell" = "#FFB300"))
  scale_colour_manual(values = c("T-PLL" = "#64B5F6", "T-NHL" = "#A5D6A7")) +
  scale_fill_manual(values = c("T-PLL" = "#64B5F6", "T-NHL" = "#A5D6A7"))
p3  

ggsave(plot = p3, filename = paste0(opt$plot, "tpll_other_t_biri.png"), 
       height = 5, width = 7)

```

#### Heatmap drug-drug correlation
```{r fig.height= 12.5, fig.width=12.5}
screenDat <- t_lymphoma

## Compute average drug effect for each patient
screenDat <- screenDat %>% 
  separate(name, into = c("col1", "col2"), sep = "[|]") %>%
  mutate(col2 = str_replace(col2, pattern = "Compound 26 \\(0.1\\)", replacement = "Compound 26")) %>%
    mutate(name = ifelse(!is.na(col2), paste0(col1, "|", col2), col1)) %>%
  select(-col1, -col2) %>%
  mutate(name = ifelse(name == "QVD-Oph", "Q-VD-OPh", name))

viabMat <- screenDat %>% filter(!diagnosis %in% c("CLL", "MCL", "MZL", "DLBCL"), 
                                 !name %in% c(drugComb)) %>%
  filter(!c(name %in% c("Birinapant", "QVD-Oph") & screen == "T_lymphoma_combi")) %>% # use 10 concentrations from the T_lymphoma plates
#  filter(!screen == "T_lymphoma_combi") %>%
  group_by(patientID, name) %>% summarise(viab = mean(viab, na.rm = TRUE)) %>%
  spread(key = patientID, value = "viab") %>% data.frame() %>%
  column_to_rownames("name") %>% as.matrix()

## Overlay drug response 
viabMat <- viabMat[complete.cases(viabMat),]

## Define colour scheme
colorList <- c(colorRampPalette(c("lightseagreen", "white"))(20),
               colorRampPalette(c("white"))(10),
               colorRampPalette(c("white","deeppink"))(20))

# corMat <- cor(t(viabMat), method = "spearman")
corMat <- cor(t(viabMat), method = "pearson")

# p4 <- pheatmap(corMat, color = colorList, 
#                breaks = seq(-1,1,length.out = 50), clustering_method = "ward.D2", name = "R^2",
#                treeheight_row = 0, #show_colnames = FALSE, 
#                row_names_gp = gpar(col = c(rep("red", 10), rep("blue", 8))),
#                main = "Drug-Drug Correlation Across All T-cell Entities")
# p4

## Get position of drugs to highlight
drugHigh <- c("Birinapant", "GDC-0152", "Nutlin-3a", "Selinexor", "Bafilomycin A1", "Dacinostat")

## Assemble colour in a vector
drugCol <- data.frame(name = rownames(corMat), col = rep("black", nrow(corMat))) %>%
  mutate(col = ifelse(name %in% drugHigh, "#D32F2F", "black"))


p4 <- Heatmap(corMat, 
        clustering_method_columns = "ward.D2", 
        clustering_method_rows = "ward.D2",
        row_names_gp = gpar(col = drugCol$col), 
        name = "R^2", border = TRUE,
        column_title = "Drug-Drug Correlation Across All T-cell Entities", 
        col=colorList, #colorRamp2(c(-1, 0, 1), colors = c("lightseagreen", "white", "deeppink")),
        column_names_gp = gpar(col = drugCol$col), 
        rect_gp = gpar(col = "black", lwd = 0.25))
p4
png(file=paste0(opt$plot, "T-lymph-Drug-Drug.png"), 
    height = 25, width = 25, res = 600, units = "cm", bg = "transparent")
draw(p4, background = "transparent")
dev.off()

```

#### Scatter plots drug-drug correlation - Birinapant
```{r fig.height= 25, fig.width=25, warning=FALSE, message=FALSE}
screenDat <- t_lymphoma

#### Scatter plots of individual drugs.
screenData <- screenDat %>% filter(!c(name %in% c("Birinapant", "QVD-Oph") & screen == "T_lymphoma_combi")) %>% # use 10 concentrations from the T_lymphoma plates
  select(patientID, diagnosis, name, viab) %>% unique() %>%
  mutate(name = str_replace(name,"Compound 26", "Compound_26")) %>%
  mutate(name = str_replace(name,"Bafilomycin A1", "Bafilomycin_A1")) %>%
  filter(diagnosis %in% c("T-PLL", "T-LGL", "AITL", "Sezary", "PTCL"), 
         !name %in% drugComb) #other diseases don't have enough samples

screenData <- screenData %>%
  mutate(diagnosis = ifelse(diagnosis %in% c("PTCL", "AITL", "T-LGL", "Sezary"), "T-NHL", diagnosis))

compDrug1 <- c("Birinapant")
compDrug2.selec <- c("Venetoclax", "Fludarabine", "Bafilomycin_A1", "Nutlin-3a", "Selinexor", "Dacinostat", "GDC-0152", "Motolimod", "AT9283", "Dasatinib", "Ruxolitinib", "Tofacitinib")
compDrug2 <- screenData %>% filter(name != compDrug1) %>% pull(name) %>% unique()

testTab <- screenData %>% 
  group_by(patientID, name) %>% mutate(viab.auc = mean(viab, na.rm = TRUE)) %>%
  select(patientID, diagnosis, name, viab.auc) %>% unique() %>%
  filter(name %in% c(compDrug1, compDrug2)) %>%
  pivot_wider(names_from = "name", values_from = viab.auc) %>%
  pivot_longer(cols = compDrug1, names_to = "drug1", values_to = "auc1") %>%
  pivot_longer(cols = compDrug2, names_to = "drug2", values_to = "auc2")

## Make plots 
pList <- lapply(compDrug2.selec, function(x) {
  p <- testTab %>% filter(drug2 == x) %>%
    ggplot(aes(x = auc1, y = auc2)) +
    geom_smooth(method = "lm", colour = "black", fill = "lightgrey") +
    geom_point(shape = 21, size = 4, aes(fill = diagnosis)) +
    xlab("Avg. Viability Birinapant") + ylab(paste0("Avg. Viability ", x)) +
    theme_classic() +
    theme(text = element_text(size = 17.5), 
          axis.text = element_text(size = 16),
          legend.title = element_blank(),
          axis.line = element_line(linewidth = 0.75),
          legend.key = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(fill = "transparent",colour = NA),
          plot.background = element_rect(fill = "transparent",colour = NA),
          legend.background = element_rect(fill='transparent')) +
    scale_fill_manual(values = c("T-PLL" = "#64B5F6", "T-NHL" = "#A5D6A7"))
  p
})
names(pList) <- compDrug2.selec
grid.arrange(grobs = pList)


lapply(compDrug2.selec, function(x) {
   p <- pList[[x]]
   ggsave(plot = p, filename = paste0(opt$plot, "corr_", x, ".png"), 
          height = 4.25, width = 5.5)
   message("Plot saved")
})

## Compute p-values and R2
resTab <- group_by(testTab, drug1, drug2) %>% nest() %>%
  mutate(m = map(data, ~cor.test(~auc1+auc2,.))) %>%
  mutate(res = map(m, broom::tidy)) %>%
  unnest(res) %>% ungroup() %>%
  select(drug1, drug2, estimate, p.value ) %>%
  arrange(p.value) %>%
  mutate(p.adj = p.adjust(p.value, method = "BH"))

resTab %>% filter(drug2 %in% c("Venetoclax", "Nutlin-3a", "Dacinostat", "Motolimod", "AT9283", 
                               "Dasatinib", "GDC-0152", "Ruxolitinib", "Tofacitinib", "Navitoclax",
                               "Fludarabine"))


```

#### Output session info
```{r pressure, echo=FALSE}
sessionInfo()
```
